---
title: "Allen Human Brain Atlas RNA-seq data processing & exploration"
output: html_notebook
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## load packages

options(warn = - 1)

```{r, message=FALSE, warning=FALSE}
library(openxlsx)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(scatterplot3d)
library(limma)
library(sva)
library(ComplexHeatmap)
library(reshape2)
library(circlize)

source("./R/function_ALLENexpdata_exploration.R")
outPath <- "./results"
```

## load Allen expression data

```{r}
# load AD vs Norm data

ALLENexp.ADNorm.col<-read.csv("./data/ALLENdata_disease_exploration/ALLENexp_ADvsNorm_columnssamples.csv")
rownames(ALLENexp.ADNorm.col)<-ALLENexp.ADNorm.col[["rnaseq_profile_id"]]
ALLENexp.ADNorm<-read.csv("./data/ALLENdata_disease_exploration/ALLENexp_ADvsNorm_log2fpkmplus1_normalized.csv")
rownames(ALLENexp.ADNorm)<-ALLENexp.ADNorm[["gene_entrez_id"]];ALLENexp.ADNorm %<>% select(-c("gene_entrez_id"))
ALLENexp.ADNorm_rowgenes<-read.csv("./data/ALLENdata_disease_exploration/ALLENexp_ADvsNorm_rowsgenes.csv")
rownames(ALLENexp.ADNorm_rowgenes)<-ALLENexp.ADNorm_rowgenes[["gene_entrez_id"]]

ALLENexp.ADNorm.col[which(ALLENexp.ADNorm.col$structure_acronym=="HIP"),"structure_color"]="#339900"
ALLENexp.ADNorm.col[which(ALLENexp.ADNorm.col$structure_acronym=="FWM"),"structure_color"]="#0066CC"
ALLENexp.ADNorm.col[which(ALLENexp.ADNorm.col$structure_acronym=="TCx"),"structure_color"]="#6600CC"
ALLENexp.ADNorm.col[which(ALLENexp.ADNorm.col$structure_acronym=="PCx"),"structure_color"]="#FF9933"

sort(unique(ALLENexp.ADNorm.col$age))

# load Dementia vs Norm data

ALLENexp.col<-read.csv("./data/ALLENdata_disease_exploration/ALLENexp_columnssamples.csv")
rownames(ALLENexp.col)<-ALLENexp.col[["rnaseq_profile_id"]]
ALLENexp<-read.csv("./data/ALLENdata_disease_exploration/ALLENexp_log2fpkmplus1_normalized.csv")
rownames(ALLENexp)<-ALLENexp[["gene_entrez_id"]];ALLENexp %<>% select(-c("gene_entrez_id"))
ALLENexp.rowgenes<-read.csv("./data/ALLENdata_disease_exploration/ALLENexp_rowsgenes.csv")
rownames(ALLENexp.rowgenes)<-ALLENexp.rowgenes[["gene_entrez_id"]]

ALLENexp.col[which(ALLENexp.col$structure_acronym=="HIP"),"structure_color"]="#339900" #green HIP
ALLENexp.col[which(ALLENexp.col$structure_acronym=="FWM"),"structure_color"]="#0066CC" #blue FWM
ALLENexp.col[which(ALLENexp.col$structure_acronym=="TCx"),"structure_color"]="#6600CC" #purple TCx
ALLENexp.col[which(ALLENexp.col$structure_acronym=="PCx"),"structure_color"]="#FF9933" #orange PCx
```

################################################################################
#
# Section 1: AD vs Norm
#
################################################################################

## PCA analysis and de-outlier

### outlier detection

```{r, message=FALSE, warning=FALSE}
probesetvar = apply(ALLENexp.ADNorm, 1, var)
ord = order(probesetvar, decreasing=TRUE)
pca = prcomp(t(ALLENexp.ADNorm[ord,]), scale=TRUE)
ss=summary(pca)
pca_table<-pca$x %>% as.data.frame %>% apply(.,2,as.numeric)
rownames(pca_table)<-colnames(ALLENexp.ADNorm)
var_explained <- pca$sdev^2/sum(pca$sdev^2)

var_explained[1:85]
sum(var_explained[1:15]) # 0.3845332

outliers=apply(pca_table[,1:15],2,function(x)which(abs(x-mean(x))>6*sd(x)))

outliers
```

### scatter plot & screeplot

```{r, message=FALSE, warning=FALSE}
# sample scatter plot

par(mfrow = c( 1, 1 ))
plot(pca$x[,1:2],type="p",xlab=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
     ylab=paste0("PC2: ",round(var_explained[2]*100,1),"%"),
     col=ALLENexp.ADNorm.col$structure_color,pch=21,main="sample PCA figure",cex.main=1)
text(pca$x,labels = rownames(pca$x),pos = 1,offset = 0.6,cex = 0.3)

scatterplot3d(pca$x[,1:3],pch=c(20),color = ALLENexp.ADNorm.col$structure_color)

# PCA screeplot

screeplot(pca, npcs = 198, type = "lines")
```

## AD vs Norm: differentially expressed analysis

### data preprocessing

```{r, message=FALSE, warning=FALSE}
ALLENexp.ADNorm.col_FWM<-ALLENexp.ADNorm.col %>% filter(structure_acronym=="FWM")

AD_id<-ALLENexp.ADNorm.col_FWM %>% filter(dsm_iv_clinical_diagnosis=="Alzheimers Disease Type") %>% .$rnaseq_profile_id %>% as.character %>% paste("X",.,sep = '')
nodementia_id<-ALLENexp.ADNorm.col_FWM %>% filter(dsm_iv_clinical_diagnosis=="No Dementia") %>% .$rnaseq_profile_id %>% as.character %>% paste("X",.,sep = '')

AD_id_exp<-ALLENexp.ADNorm %>% select(all_of(AD_id))
nodementia_id_exp<-ALLENexp.ADNorm %>% select(all_of(nodementia_id))
```

### example: single-gene level DE analysis

```{r}
shapiro.test(as.numeric(AD_id_exp[which(rownames(AD_id_exp)==80217),]))
shapiro.test(as.numeric(nodementia_id_exp[which(rownames(AD_id_exp)==80217),]))
t.test(AD_id_exp[which(rownames(AD_id_exp)==80217),],nodementia_id_exp[which(rownames(nodementia_id_exp)==80217),])
group=c(rep("AD",length(AD_id_exp[which(rownames(AD_id_exp)==80217),])),rep("nodementia",length(nodementia_id_exp[which(rownames(nodementia_id_exp)==80217),])))
gene_exp=c(as.numeric(AD_id_exp[which(rownames(AD_id_exp)==80217),]),as.numeric(nodementia_id_exp[which(rownames(nodementia_id_exp)==80217),]))
df_exp=as.data.frame(cbind(group,gene_exp))
df_exp$gene_exp=unlist(lapply(df_exp$gene_exp,as.numeric))

p <- ggboxplot(df_exp, x = "group", y = "gene_exp",
               color = "group", palette = "jco",
               add = "jitter")
p + stat_compare_means(aes(group = group),method = "t.test",label = "p.signif")

p <- ggviolin(df_exp,x="group",y="gene_exp",width=0.7,size=0.7,
              fill="group",palette = c("#00AFBB", "#E7B800", "#FC4E07"),
              add = "boxplot", add.params = list(fill = "white"),shape="group",x.text.angle = 0,
              axis.text.x=element_text(color="black",size=rel(1.1)),
              axis.text.y=element_text(color="black", size=rel(1.1)),
              axis.title.x = element_text(color="black", size=rel(1.1))
)

p

p + #stat_compare_means(comparisons = my_comparisons)+#,method='t.test') +
  stat_compare_means(label.y = 9.5)#,method='t.test')

p + stat_compare_means(method = "t.test")
```

### load testing genesets

```{r}
# burdened genes
burdengene_AD<-read.csv("./metadata/list_burdengenes_AD.csv")
burdengene_ADearly<-read.csv("./metadata/list_burdengenes_ADearly.csv")
burdengene_dementia<-read.csv("./metadata/list_burdengenes_dementia.csv")

# case-only and burden genes
list_caseonly_dementia<-read.table("./metadata/list_TRAPD_caseonly_dementia_entrezid.txt")[,1]
list_caseonly_AD<-read.table("./metadata/list_TRAPD_caseonly_AD_entrezid.txt")[,1]
list_caseonly_earlyAD<-read.table("./metadata/list_TRAPD_caseonly_earlyAD_entrezid.txt")[,1]
list_caseonly_and_burden_all<-read.table("./metadata/list_TRAPD_caseonly_mutatedgenes_an_burdengenes.txt",header = T)
```

### DE analysis execution

```{r, message=FALSE, warning=FALSE}
# burdened genes
FWM_combat_exp<-get_exprmat_ALLEN(ALLENexp.ADNorm,ALLENexp.ADNorm.col,"FWM","AD",burdengene_dementia$ENTREZID)
TCx_combat_exp<-get_exprmat_ALLEN(ALLENexp.ADNorm,ALLENexp.ADNorm.col,"TCx","AD",burdengene_dementia$ENTREZID)
HIP_combat_exp<-get_exprmat_ALLEN(ALLENexp.ADNorm,ALLENexp.ADNorm.col,"HIP","AD",burdengene_dementia$ENTREZID)
PCx_combat_exp<-get_exprmat_ALLEN(ALLENexp.ADNorm,ALLENexp.ADNorm.col,"PCx","AD",burdengene_dementia$ENTREZID)

# case-only and burden genes
AD_caseonlyburden_FWM_combat_exp<-get_exprmat_ALLEN(ALLENexp.ADNorm,ALLENexp.ADNorm.col,"FWM","AD",list_caseonly_and_burden_all$ENTREZID)
AD_caseonlyburden_TCx_combat_exp<-get_exprmat_ALLEN(ALLENexp.ADNorm,ALLENexp.ADNorm.col,"TCx","AD",list_caseonly_and_burden_all$ENTREZID)
AD_caseonlyburden_HIP_combat_exp<-get_exprmat_ALLEN(ALLENexp.ADNorm,ALLENexp.ADNorm.col,"HIP","AD",list_caseonly_and_burden_all$ENTREZID)
AD_caseonlyburden_PCx_combat_exp<-get_exprmat_ALLEN(ALLENexp.ADNorm,ALLENexp.ADNorm.col,"PCx","AD",list_caseonly_and_burden_all$ENTREZID)
```

### example: normal distribution testing

```{r}
shapiro.test(FWM_combat_exp[1,])
shapiro.test(PCx_combat_exp[10,])
shapiro.test(HIP_combat_exp[10,])
shapiro.test(TCx_combat_exp[1,])

hist(TCx_combat_exp[9,])
```

## z-score transformation & mark significantly DE genes

```{r, message=FALSE, warning=FALSE}
# burdened genes
zcore_signifgenes_FWM=get_signifgenes(FWM_combat_exp,"FWM","AD",fdr_threshold=0.05)
zcore_signifgenes_TCx=get_signifgenes(TCx_combat_exp,"TCx","AD",fdr_threshold=0.05)
zcore_signifgenes_HIP=get_signifgenes(HIP_combat_exp,"HIP","AD",fdr_threshold=0.05)
zcore_signifgenes_PCx=get_signifgenes(PCx_combat_exp,"PCx","AD",fdr_threshold=0.05)

# case-only and burden genes
AD_caseonlyburden_zcore_signifgenes_FWM=get_signifgenes(AD_caseonlyburden_FWM_combat_exp,"FWM","AD",fdr_threshold=0.05)
AD_caseonlyburden_zcore_signifgenes_TCx=get_signifgenes(AD_caseonlyburden_TCx_combat_exp,"TCx","AD",fdr_threshold=0.05)
AD_caseonlyburden_zcore_signifgenes_HIP=get_signifgenes(AD_caseonlyburden_HIP_combat_exp,"HIP","AD",fdr_threshold=0.05)
AD_caseonlyburden_zcore_signifgenes_PCx=get_signifgenes(AD_caseonlyburden_PCx_combat_exp,"PCx","AD",fdr_threshold=0.05)
```

### output significantly DE gene results

```{r}
DE_list.AD<-list()
DE_list.AD$burdened.FWM<-zcore_signifgenes_FWM
DE_list.AD$burdened.TCx<-zcore_signifgenes_TCx
DE_list.AD$burdened.HIP<-zcore_signifgenes_HIP
DE_list.AD$burdened.PCx<-zcore_signifgenes_PCx
DE_list.AD$caseonlyburden.FWM<-AD_caseonlyburden_zcore_signifgenes_FWM
DE_list.AD$caseonlyburden.TCx<-AD_caseonlyburden_zcore_signifgenes_TCx
DE_list.AD$caseonlyburden.HIP<-AD_caseonlyburden_zcore_signifgenes_HIP
DE_list.AD$caseonlyburden.PCx<-AD_caseonlyburden_zcore_signifgenes_PCx
write.xlsx(DE_list.AD,paste(outPath,"z-score_sigDE_fdr05-adjusted_ADvsNorm_genes.xlsx",sep = '/'))
```

## Plot drawing

### heatmap plot

```{r, message=FALSE, warning=FALSE}
# burdened genes

draw_heatmap_zscore(zcore_signifgenes_FWM,"AD","FWM",fdr_threshold = 0.05)
draw_heatmap_zscore(zcore_signifgenes_TCx,"AD","TCx",fdr_threshold = 0.05)
draw_heatmap_zscore(zcore_signifgenes_HIP,"AD","HIP",fdr_threshold = 0.05)
draw_heatmap_zscore(zcore_signifgenes_PCx,"AD","PCx",fdr_threshold = 0.05)

# case-only and burden genes

draw_heatmap_zscore(AD_caseonlyburden_zcore_signifgenes_FWM,"AD","FWM",fdr_threshold=0.05)
draw_heatmap_zscore(AD_caseonlyburden_zcore_signifgenes_TCx,"AD","TCx",fdr_threshold=0.05)
draw_heatmap_zscore(AD_caseonlyburden_zcore_signifgenes_HIP,"AD","HIP",fdr_threshold=0.05)
draw_heatmap_zscore(AD_caseonlyburden_zcore_signifgenes_PCx,"AD","PCx",fdr_threshold=0.05)
```

### violin plot

```{r, message=FALSE, warning=FALSE}
# burdened genes

draw_violin_plots(zcore_signifgenes_FWM,"AD","FWM",fdr_threshold=0.05)
draw_violin_plots(zcore_signifgenes_TCx,"AD","TCx",fdr_threshold=0.05)
draw_violin_plots(zcore_signifgenes_HIP,"AD","HIP",fdr_threshold=0.05)
draw_violin_plots(zcore_signifgenes_PCx,"AD","PCx",fdr_threshold=0.05)

# case-only and burden genes

draw_violin_plots(AD_caseonlyburden_zcore_signifgenes_FWM,"AD","FWM",fdr_threshold=0.05)
draw_violin_plots(AD_caseonlyburden_zcore_signifgenes_TCx,"AD","TCx",fdr_threshold=0.05)
draw_violin_plots(AD_caseonlyburden_zcore_signifgenes_HIP,"AD","HIP",fdr_threshold=0.05)
draw_violin_plots(AD_caseonlyburden_zcore_signifgenes_PCx,"AD","PCx",fdr_threshold=0.05)
```

################################################################################
#
# Section 2: Dementia vs Norm
#
################################################################################

# differential expression analysis

```{r, message=FALSE, warning=FALSE}
# burdened genes

Dementia_FWM_combat_exp<-get_exprmat_ALLEN(ALLENexp,ALLENexp.col,"FWM","dementia",burdengene_dementia$ENTREZID)
Dementia_TCx_combat_exp<-get_exprmat_ALLEN(ALLENexp,ALLENexp.col,"TCx","dementia",burdengene_dementia$ENTREZID)
Dementia_HIP_combat_exp<-get_exprmat_ALLEN(ALLENexp,ALLENexp.col,"HIP","dementia",burdengene_dementia$ENTREZID)
Dementia_PCx_combat_exp<-get_exprmat_ALLEN(ALLENexp,ALLENexp.col,"PCx","dementia",burdengene_dementia$ENTREZID)

# case-only and burden genes

Dementia_caseonlyburden_FWM_combat_exp<-get_exprmat_ALLEN(ALLENexp,ALLENexp.col,"FWM","dementia",list_caseonly_and_burden_all$ENTREZID)
Dementia_caseonlyburden_TCx_combat_exp<-get_exprmat_ALLEN(ALLENexp,ALLENexp.col,"TCx","dementia",list_caseonly_and_burden_all$ENTREZID)
Dementia_caseonlyburden_HIP_combat_exp<-get_exprmat_ALLEN(ALLENexp,ALLENexp.col,"HIP","dementia",list_caseonly_and_burden_all$ENTREZID)
Dementia_caseonlyburden_PCx_combat_exp<-get_exprmat_ALLEN(ALLENexp,ALLENexp.col,"PCx","dementia",list_caseonly_and_burden_all$ENTREZID)
```

## z-score transformation & mark significantly DE genes

```{r, message=FALSE, warning=FALSE}
# burdened genes

Dementia_zcore_signifgenes_FWM=get_signifgenes(Dementia_FWM_combat_exp,"FWM","dementia",fdr_threshold=0.05)
Dementia_zcore_signifgenes_TCx=get_signifgenes(Dementia_TCx_combat_exp,"TCx","dementia",fdr_threshold=0.05)
Dementia_zcore_signifgenes_HIP=get_signifgenes(Dementia_HIP_combat_exp,"HIP","dementia",fdr_threshold=0.05)
Dementia_zcore_signifgenes_PCx=get_signifgenes(Dementia_PCx_combat_exp,"PCx","dementia",fdr_threshold=0.05)

# case-only and burden genes

Dementia_caseonlyburden_zcore_signifgenes_FWM=get_signifgenes(Dementia_caseonlyburden_FWM_combat_exp,"FWM","dementia",fdr_threshold=0.05)
Dementia_caseonlyburden_zcore_signifgenes_TCx=get_signifgenes(Dementia_caseonlyburden_TCx_combat_exp,"TCx","dementia",fdr_threshold=0.05)
Dementia_caseonlyburden_zcore_signifgenes_HIP=get_signifgenes(Dementia_caseonlyburden_HIP_combat_exp,"HIP","dementia",fdr_threshold=0.05)
Dementia_caseonlyburden_zcore_signifgenes_PCx=get_signifgenes(Dementia_caseonlyburden_PCx_combat_exp,"PCx","dementia",fdr_threshold=0.05)
```

### output significantly DE gene results

```{r}
DE_list.Dementia<-list()
DE_list.Dementia$burdened.FWM<-Dementia_zcore_signifgenes_FWM
DE_list.Dementia$burdened.TCx<-Dementia_zcore_signifgenes_TCx
DE_list.Dementia$burdened.HIP<-Dementia_zcore_signifgenes_HIP
DE_list.Dementia$burdened.PCx<-Dementia_zcore_signifgenes_PCx
DE_list.Dementia$caseonlyburden.FWM<-Dementia_caseonlyburden_zcore_signifgenes_FWM
DE_list.Dementia$caseonlyburden.TCx<-Dementia_caseonlyburden_zcore_signifgenes_TCx
DE_list.Dementia$caseonlyburden.HIP<-Dementia_caseonlyburden_zcore_signifgenes_HIP
DE_list.Dementia$caseonlyburden.PCx<-Dementia_caseonlyburden_zcore_signifgenes_PCx
write.xlsx(DE_list.Dementia,paste(outPath,"z-score_sigDE_fdr05-adjusted_DementiavsNorm_genes.xlsx",sep = '/'))
```

## Plot drawing

### heatmap plot

```{r, message=FALSE, warning=FALSE}
# burdened genes

draw_heatmap_zscore(Dementia_zcore_signifgenes_TCx,"dementia","TCx",fdr_threshold = 0.05)
draw_heatmap_zscore(Dementia_zcore_signifgenes_HIP,"dementia","HIP",fdr_threshold = 0.05)
draw_heatmap_zscore(Dementia_zcore_signifgenes_PCx,"dementia","PCx",fdr_threshold = 0.05)

# case-only and burden genes

draw_heatmap_zscore(Dementia_caseonlyburden_zcore_signifgenes_FWM,"dementia","FWM",fdr_threshold=0.05)
draw_heatmap_zscore(Dementia_caseonlyburden_zcore_signifgenes_TCx,"dementia","TCx",fdr_threshold=0.05)
draw_heatmap_zscore(Dementia_caseonlyburden_zcore_signifgenes_HIP,"dementia","HIP",fdr_threshold=0.05)
```

### violin plot

```{r, message=FALSE, warning=FALSE}
# burdened genes

draw_violin_plots(Dementia_zcore_signifgenes_TCx,"dementia","TCx",fdr_threshold=0.05)
draw_violin_plots(Dementia_zcore_signifgenes_HIP,"dementia","HIP",fdr_threshold=0.05)
draw_violin_plots(Dementia_zcore_signifgenes_PCx,"dementia","PCx",fdr_threshold=0.05)

# case-only and burden genes

draw_violin_plots(Dementia_caseonlyburden_zcore_signifgenes_FWM,"dementia","FWM",fdr_threshold=0.05)
draw_violin_plots(Dementia_caseonlyburden_zcore_signifgenes_TCx,"dementia","TCx",fdr_threshold=0.05)
draw_violin_plots(Dementia_caseonlyburden_zcore_signifgenes_HIP,"dementia","HIP",fdr_threshold=0.05)
```

## add evidence to the ranking list

### extract DE gene lists

```{r}
# the previous script in "singlecell_marker_enrichment_analysis_CellMarker"

## AD vs Norm

### burdened genes

gene_df<-data.frame()

tmp_df<-cbind(rownames(zcore_signifgenes_FWM),zcore_signifgenes_FWM$padjust_fdr,zcore_signifgenes_FWM$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(zcore_signifgenes_TCx),zcore_signifgenes_TCx$padjust_fdr,zcore_signifgenes_TCx$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(zcore_signifgenes_HIP),zcore_signifgenes_HIP$padjust_fdr,zcore_signifgenes_HIP$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(zcore_signifgenes_PCx),zcore_signifgenes_PCx$padjust_fdr,zcore_signifgenes_PCx$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

siggenes.ADvsNorm.burdened<-table(gene_df$symbol) %>% as.data.frame %>% arrange(desc(Freq)); colnames(siggenes.ADvsNorm.burdened)<-c("symbol","Freq")
siggenes.ADvsNorm.burdened$EntrezID<-siggenes.ADvsNorm.burdened$symbol %>% lapply(.,function(x){ID_conversion_tr[which(ID_conversion_tr$SYMBOL==x),"ENTREZID"]}) %>% unlist

### case-only and burdened genes

gene_df<-data.frame()

tmp_df<-cbind(rownames(AD_caseonlyburden_zcore_signifgenes_FWM),AD_caseonlyburden_zcore_signifgenes_FWM$padjust_fdr,AD_caseonlyburden_zcore_signifgenes_FWM$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(AD_caseonlyburden_zcore_signifgenes_TCx),AD_caseonlyburden_zcore_signifgenes_TCx$padjust_fdr,AD_caseonlyburden_zcore_signifgenes_TCx$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(AD_caseonlyburden_zcore_signifgenes_HIP),AD_caseonlyburden_zcore_signifgenes_HIP$padjust_fdr,AD_caseonlyburden_zcore_signifgenes_HIP$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(AD_caseonlyburden_zcore_signifgenes_PCx),AD_caseonlyburden_zcore_signifgenes_PCx$padjust_fdr,AD_caseonlyburden_zcore_signifgenes_PCx$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

siggenes.ADvsNorm.caseonlyburden<-table(gene_df$symbol) %>% as.data.frame %>% arrange(desc(Freq)); colnames(siggenes.ADvsNorm.caseonlyburden)<-c("symbol","Freq")
siggenes.ADvsNorm.caseonlyburden$EntrezID<-siggenes.ADvsNorm.caseonlyburden$symbol %>% lapply(.,function(x){ID_conversion_tr[which(ID_conversion_tr$SYMBOL==x),"ENTREZID"]}) %>% unlist

## dementia vs Norm

### burdened genes

gene_df<-data.frame()

#### "no significantly DE genes" for "Dementia_zcore_signifgenes_FWM"
tmp_df<-cbind(rownames(Dementia_zcore_signifgenes_TCx),Dementia_zcore_signifgenes_TCx$padjust_fdr,Dementia_zcore_signifgenes_TCx$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(Dementia_zcore_signifgenes_HIP),Dementia_zcore_signifgenes_HIP$padjust_fdr,Dementia_zcore_signifgenes_HIP$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(Dementia_zcore_signifgenes_PCx),Dementia_zcore_signifgenes_PCx$padjust_fdr,Dementia_zcore_signifgenes_PCx$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

siggenes.DementiavsNorm.burdened<-table(gene_df$symbol) %>% as.data.frame %>% arrange(desc(Freq)); colnames(siggenes.DementiavsNorm.burdened)<-c("symbol","Freq")
siggenes.DementiavsNorm.burdened$EntrezID<-siggenes.DementiavsNorm.burdened$symbol %>% lapply(.,function(x){ID_conversion_tr[which(ID_conversion_tr$SYMBOL==x),"ENTREZID"]}) %>% unlist

### case-only and burdened genes

gene_df<-data.frame()

tmp_df<-cbind(rownames(Dementia_caseonlyburden_zcore_signifgenes_FWM),Dementia_caseonlyburden_zcore_signifgenes_FWM$padjust_fdr,Dementia_caseonlyburden_zcore_signifgenes_FWM$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(Dementia_caseonlyburden_zcore_signifgenes_TCx),Dementia_caseonlyburden_zcore_signifgenes_TCx$padjust_fdr,Dementia_caseonlyburden_zcore_signifgenes_TCx$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)

tmp_df<-cbind(rownames(Dementia_caseonlyburden_zcore_signifgenes_HIP),Dementia_caseonlyburden_zcore_signifgenes_HIP$padjust_fdr,Dementia_caseonlyburden_zcore_signifgenes_HIP$sig_DE) %>% as.data.frame; colnames(tmp_df)<-c("symbol","padjust_fdr","sig_DE")
tmp_df %<>% dplyr::filter(sig_DE=="TRUE")
gene_df %<>% rbind(.,tmp_df)
#### "no significantly DE genes" for "Dementia_caseonlyburden_zcore_signifgenes_PCx"

siggenes.DementiavsNorm.caseonlyburden<-table(gene_df$symbol) %>% as.data.frame %>% arrange(desc(Freq)); colnames(siggenes.DementiavsNorm.caseonlyburden)<-c("symbol","Freq")
siggenes.DementiavsNorm.caseonlyburden$EntrezID<-siggenes.DementiavsNorm.caseonlyburden$symbol %>% lapply(.,function(x){ID_conversion_tr[which(ID_conversion_tr$SYMBOL==x),"ENTREZID"]}) %>% unlist
```

### output the result lists

```{r}
out_DE.list<-list()
out_DE.list$ADvsNorm_burdened<-siggenes.ADvsNorm.burdened
out_DE.list$ADvsNorm_caseonlyburden<-siggenes.ADvsNorm.caseonlyburden
out_DE.list$DementiavsNorm_burdened<-siggenes.DementiavsNorm.burdened
out_DE.list$DementiavsNorm_caseonlyburden<-siggenes.DementiavsNorm.caseonlyburden
write.xlsx(out_DE.list,paste(outPath,"sig_DE_gene_tables_allgroups_allgenesets.xlsx",sep = '/'))
```

### add evidence

```{r}
evidence_df<-read.xlsx("./metadata/evidence-ranking_table_all-case-only-and-burdened-genes.xlsx")

evidence_df$ADvsNorm_caseonlyburden_DE<-evidence_df$ENTREZID %>% lapply(.,function(x){if(x %in% siggenes.ADvsNorm.caseonlyburden$EntrezID){siggenes.ADvsNorm.caseonlyburden[which(siggenes.ADvsNorm.caseonlyburden$EntrezID==x),"Freq"]}else{0}}) %>% unlist
evidence_df$DementiavsNorm_caseonlyburden_DE<-evidence_df$ENTREZID %>% lapply(.,function(x){if(x %in% siggenes.DementiavsNorm.caseonlyburden$EntrezID){siggenes.DementiavsNorm.caseonlyburden[which(siggenes.DementiavsNorm.caseonlyburden$EntrezID==x),"Freq"]}else{0}}) %>% unlist

write.xlsx(evidence_df,"./metadata/evidence-ranking_table_all-case-only-and-burdened-genes.xlsx")

# the next script is "CNV result filtering"
```

## save.image

```{r}
save.image("./ALLENexpdata_exploration.RData")
```
