---
title: "single-cell enrichment analysis and figure drawing"
output: html_notebook
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## load packages

```{r, message=FALSE, warning=FALSE}
library(openxlsx)
library(clusterProfiler)
library(org.Hs.eg.db)
library(biomaRt)
library(lattice)
library(gplots)
library(RColorBrewer)
library(magrittr)
library(tidyverse)

source("./R/function_singlecell_marker_enrichment_analysis_CellMarker.R")
outPath <- "./results"
```

################################################################################
#
# Background table preparation
#
################################################################################

## load original mutation table

```{r, message=FALSE, warning=FALSE}
burdenresult_dementia<-read.xlsx("./data/burdengenes_result_extraction.xlsx",sheet = "dementia")
burdenresult_AD<-read.xlsx("./data/burdengenes_result_extraction.xlsx",sheet = "AD")
burdenresult_earlyAD<-read.xlsx("./data/burdengenes_result_extraction.xlsx",sheet = "ADearly")
burdenresult_dementia$disease<-rep(0,length(burdenresult_dementia$GENE))
burdenresult_AD$disease<-rep(0,length(burdenresult_AD$GENE))
burdenresult_earlyAD$disease<-rep(0,length(burdenresult_earlyAD$GENE))
```

## Entrez ID conversion

```{r, message=FALSE, warning=FALSE}
burdenresult_dementia.entrez_id<-bitr(burdenresult_dementia$GENE, fromType = "SYMBOL",
                                      toType = c("ENTREZID"),
                                      OrgDb = org.Hs.eg.db) %>% dplyr::rename(c(GENE=SYMBOL, entrezgene_id=ENTREZID))

burdenresult_AD.entrez_id<-bitr(burdenresult_AD$GENE, fromType = "SYMBOL",
                                      toType = c("ENTREZID"),
                                      OrgDb = org.Hs.eg.db) %>% dplyr::rename(c(GENE=SYMBOL, entrezgene_id=ENTREZID))

burdenresult_earlyAD.entrez_id<-bitr(burdenresult_earlyAD$GENE, fromType = "SYMBOL",
                                      toType = c("ENTREZID"),
                                      OrgDb = org.Hs.eg.db) %>% dplyr::rename(c(GENE=SYMBOL, entrezgene_id=ENTREZID))

burdenresult_dementia<-merge(burdenresult_dementia,burdenresult_dementia.entrez_id,by=c("GENE"))
burdenresult_AD<-merge(burdenresult_AD,burdenresult_AD.entrez_id,by=c("GENE"))
burdenresult_earlyAD<-merge(burdenresult_earlyAD,burdenresult_earlyAD.entrez_id,by=c("GENE"))

rm(burdenresult_dementia.entrez_id,burdenresult_AD.entrez_id,burdenresult_earlyAD.entrez_id)
```

## further processing the background gene tables

```{r, message=FALSE, warning=FALSE}
# Delete genes without mapped entrezgene_id (entrezgene_id==0)

burdenresult_dementia %<>% filter(entrezgene_id!=0)
burdenresult_AD %<>% filter(entrezgene_id!=0)
burdenresult_earlyAD %<>% filter(entrezgene_id!=0)

# Select useful columns

burdenresult_dementia %<>% select(GENE,entrezgene_id,disease)
burdenresult_AD %<>% select(GENE,entrezgene_id,disease)
burdenresult_earlyAD %<>% select(GENE,entrezgene_id,disease)
```

## obtain GC content and gene length of background genes

```{r, message=FALSE, warning=FALSE}
mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
mart_attribute<-listAttributes(mart)
#Transcript length (including UTRs and CDS)

# get gene length & GC content

burdenresult_dementia<-get_genelengthGCcontent(burdenresult_dementia)
burdenresult_AD<-get_genelengthGCcontent(burdenresult_AD)
burdenresult_earlyAD<-get_genelengthGCcontent(burdenresult_earlyAD)

rm(mart,mart_attribute)
```

# optional for the background tables: directly load prepared tables (the one used in this paper with more gene IDs annotated)

```{r}
burdenresult_dementia<-read.csv("./metadata/burdenresult_dementia_annotatedforLR.csv") %>% dplyr::select(c("entrez_id","ensembl_gene_id","GENE","transcript_length","percentage_gene_gc_content","P_DOM_adjustFDR"))
burdenresult_AD<-read.csv("./metadata/burdenresult_AD_annotatedforLR.csv") %>% dplyr::select(c("entrez_id","ensembl_gene_id","GENE","transcript_length","percentage_gene_gc_content","P_DOM_adjustFDR"))
burdenresult_earlyAD<-read.csv("./metadata/burdenresult_earlyAD_annotatedforLR.csv") %>% dplyr::select(c("entrez_id","ensembl_gene_id","GENE","transcript_length","percentage_gene_gc_content","P_DOM_adjustFDR"))

colnames(burdenresult_dementia)<-c("entrezgene_id",colnames(burdenresult_dementia)[colnames(burdenresult_dementia) %!in% "entrez_id"])
colnames(burdenresult_AD)<-c("entrezgene_id",colnames(burdenresult_AD)[colnames(burdenresult_AD) %!in% "entrez_id"])
colnames(burdenresult_earlyAD)<-c("entrezgene_id",colnames(burdenresult_earlyAD)[colnames(burdenresult_earlyAD) %!in% "entrez_id"])

burdenresult_dementia$disease<-rep(0,nrow(burdenresult_dementia))
burdenresult_AD$disease<-rep(0,nrow(burdenresult_AD))
burdenresult_earlyAD$disease<-rep(0,nrow(burdenresult_earlyAD))
```

## testing geneset preparation

```{r}
dementia_PPI_cluster<-read.table("./metadata/dementia_PPI_cluster.txt")[,1]
AD_PPI_cluster<-read.table("./metadata/AD_PPI_cluster.txt")[,1]
earlyAD_PPI_cluster<-read.table("./metadata/ADearly_PPI_cluster.txt")[,1]

list_burden_dementia<-read.table("./metadata/burdengenes_dementia_list.txt")[,1]
list_burden_AD<-read.table("./metadata/burdengenes_AD_list.txt")[,1]
list_burden_earlyAD<-read.table("./metadata/burdengenes_ADearly_list.txt")[,1]

list_caseonly_dementia<-read.table("./metadata/list_TRAPD_caseonly_dementia_entrezid.txt")[,1]
list_caseonly_AD<-read.table("./metadata/list_TRAPD_caseonly_AD_entrezid.txt")[,1]
list_caseonly_earlyAD<-read.table("./metadata/list_TRAPD_caseonly_earlyAD_entrezid.txt")[,1]
list_caseonly_and_burden_all<-read.table("./metadata/list_TRAPD_caseonly_mutated_an_burdengenes_entrezid.txt")[,1]

list_case_AD<-read.table("./metadata/list_case_AD_entrezid.txt")[,1]
list_case_earlyAD<-read.table("./metadata/list_case_earlyAD_entrezid.txt")[,1]
list_case_dementia<-read.table("./metadata/list_case_dementia_entrezid.txt")[,1]
```

################################################################################
#
# Cell type enrichment analysis with markers from CellMarker database
#
################################################################################

## cell type marker loading and background gene annotation

```{r, message=FALSE, warning=FALSE}
celltype_markers.CellMarker<-read.table("./data/Source_CellMarker/Human_cell_markers.txt",sep = '\t',header = T)
celltype_markers.CellMarker_set<-read.table("./data/Source_CellMarker/Single_cell_markers.txt",sep = '\t',header = T)
```

# Section 1: filter human and normal samples from the marker list

## Prepare data

```{r, message=FALSE, warning=FALSE}
celltype_markers.CellMarker.hgnormal<-get_CellMarker_celltypefilter(celltype_markers.CellMarker,filtertype = c("speciesType","cancerType"),filter_value = c("Human","Normal"))
celltype_markers.CellMarker_set.hgnormal<-get_CellMarker_celltypefilter(celltype_markers.CellMarker_set,filtertype = c("speciesType","cancerType"),filter_value = c("Human","Normal"))

# cell type annotation

celltype_markers.CellMarker.hgnormal.anno_dementia<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker.hgnormal,background_genes = burdenresult_dementia)
celltype_markers.CellMarker.hgnormal.anno_AD<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker.hgnormal,background_genes = burdenresult_AD)
celltype_markers.CellMarker.hgnormal.anno_EOAD<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker.hgnormal,background_genes = burdenresult_earlyAD)

celltype_markers.CellMarker_set.hgnormal.anno_dementia<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker_set.hgnormal,background_genes = burdenresult_dementia)
celltype_markers.CellMarker_set.hgnormal.anno_AD<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker_set.hgnormal,background_genes = burdenresult_AD)
celltype_markers.CellMarker_set.hgnormal.anno_EOAD<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker_set.hgnormal,background_genes = burdenresult_earlyAD)
```

### cell-type Enrichment test from celltype_markers.CellMarker.hgnormal

```{r, message=FALSE, warning=FALSE}
CellMarker_enrich_res.burdendementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_dementia,disease_genes = list_burden_dementia)
CellMarker_enrich_res.burdenAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_AD,disease_genes = list_burden_AD)
CellMarker_enrich_res.burdenearlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_EOAD,disease_genes = list_burden_earlyAD)

CellMarker_enrich_res.caseonly_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_dementia,disease_genes = list_caseonly_dementia)
CellMarker_enrich_res.caseonly_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_AD,disease_genes = list_caseonly_AD)
CellMarker_enrich_res.caseonly_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_EOAD,disease_genes = list_caseonly_earlyAD)

CellMarker_enrich_res.case_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_dementia,disease_genes = list_case_dementia)
CellMarker_enrich_res.case_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_AD,disease_genes = list_case_AD)
CellMarker_enrich_res.case_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_EOAD,disease_genes = list_case_earlyAD)

CellMarker_enrich_res.PPI_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_dementia,disease_genes = dementia_PPI_cluster)
CellMarker_enrich_res.PPI_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_AD,disease_genes = AD_PPI_cluster)
CellMarker_enrich_res.PPI_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal.anno_EOAD,disease_genes = earlyAD_PPI_cluster)
```

### cell-type Enrichment test from celltype_markers.CellMarker_set.hgnormal

```{r, message=FALSE, warning=FALSE}
CellMarker_enrich_set_res.burdendementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_dementia,disease_genes = list_burden_dementia)
CellMarker_enrich_set_res.burdenAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_AD,disease_genes = list_burden_AD)
CellMarker_enrich_set_res.burdenearlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_EOAD,disease_genes = list_burden_earlyAD)

CellMarker_enrich_set_res.caseonly_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_dementia,disease_genes = list_caseonly_dementia)
CellMarker_enrich_set_res.caseonly_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_AD,disease_genes = list_caseonly_AD)
CellMarker_enrich_set_res.caseonly_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_EOAD,disease_genes = list_caseonly_earlyAD)

CellMarker_enrich_set_res.case_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_dementia,disease_genes = list_case_dementia)
CellMarker_enrich_set_res.case_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_AD,disease_genes = list_case_AD)
CellMarker_enrich_set_res.case_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_EOAD,disease_genes = list_case_earlyAD)

CellMarker_enrich_set_res.PPI_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_dementia,disease_genes = dementia_PPI_cluster)
CellMarker_enrich_set_res.PPI_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_AD,disease_genes = AD_PPI_cluster)
CellMarker_enrich_set_res.PPI_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal.anno_EOAD,disease_genes = earlyAD_PPI_cluster)
```

# Section 2: filter human and normal brain samples from the marker list

```{r, message=FALSE, warning=FALSE}
celltype_markers.CellMarker.hgnormal_brain<-get_CellMarker_celltypefilter(celltype_markers.CellMarker,filtertype = c("speciesType","tissueType","cancerType"),filter_value = c("Human","Brain","Normal"))
celltype_markers.CellMarker_set.hgnormal_brain<-get_CellMarker_celltypefilter(celltype_markers.CellMarker_set,filtertype = c("speciesType","tissueType","cancerType"),filter_value = c("Human","Brain","Normal"))

# cell type annnotation

celltype_markers.CellMarker.hgnormal_brain.anno_dementia<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker.hgnormal_brain,background_genes = burdenresult_dementia)
celltype_markers.CellMarker.hgnormal_brain.anno_AD<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker.hgnormal_brain,background_genes = burdenresult_AD)
celltype_markers.CellMarker.hgnormal_brain.anno_EOAD<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker.hgnormal_brain,background_genes = burdenresult_earlyAD)

celltype_markers.CellMarker_set.hgnormal_brain.anno_dementia<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker_set.hgnormal_brain,background_genes = burdenresult_dementia)
celltype_markers.CellMarker_set.hgnormal_brain.anno_AD<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker_set.hgnormal_brain,background_genes = burdenresult_AD)
celltype_markers.CellMarker_set.hgnormal_brain.anno_EOAD<-get_CellMarker_annotation_prepare(celltype_markers.CellMarker_set.hgnormal_brain,background_genes = burdenresult_earlyAD)
```

## get celltypeEnrichment from celltype_markers.CellMarker.hgnormal_brain

```{r, message=FALSE, warning=FALSE}
CellMarker_enrich_brain_res.burdendementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_dementia,disease_genes = list_burden_dementia)
CellMarker_enrich_brain_res.burdenAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_AD,disease_genes = list_burden_AD)
CellMarker_enrich_brain_res.burdenearlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_EOAD,disease_genes = list_burden_earlyAD)

CellMarker_enrich_brain_res.caseonly_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_dementia,disease_genes = list_caseonly_dementia)
CellMarker_enrich_brain_res.caseonly_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_AD,disease_genes = list_caseonly_AD)
CellMarker_enrich_brain_res.caseonly_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_EOAD,disease_genes = list_caseonly_earlyAD)

CellMarker_enrich_brain_res.case_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_dementia,disease_genes = list_case_dementia)
CellMarker_enrich_brain_res.case_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_AD,disease_genes = list_case_AD)
CellMarker_enrich_brain_res.case_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_EOAD,disease_genes = list_case_earlyAD)

CellMarker_enrich_brain_res.PPI_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_dementia,disease_genes = dementia_PPI_cluster)
CellMarker_enrich_brain_res.PPI_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_AD,disease_genes = AD_PPI_cluster)
CellMarker_enrich_brain_res.PPI_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker.hgnormal_brain.anno_EOAD,disease_genes = earlyAD_PPI_cluster)
```

## get celltypeEnrichment from celltype_markers.CellMarker_set.hgnormal_brain

```{r, message=FALSE, warning=FALSE}
CellMarker_enrich_brain_set_res.burdendementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_dementia,disease_genes = list_burden_dementia)
CellMarker_enrich_brain_set_res.burdenAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_AD,disease_genes = list_burden_AD)
CellMarker_enrich_brain_set_res.burdenearlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_EOAD,disease_genes = list_burden_earlyAD)

CellMarker_enrich_brain_set_res.caseonly_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_dementia,disease_genes = list_caseonly_dementia)
CellMarker_enrich_brain_set_res.caseonly_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_AD,disease_genes = list_caseonly_AD)
CellMarker_enrich_brain_set_res.caseonly_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_EOAD,disease_genes = list_caseonly_earlyAD)

CellMarker_enrich_brain_set_res.case_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_dementia,disease_genes = list_case_dementia)
CellMarker_enrich_brain_set_res.case_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_AD,disease_genes = list_case_AD)
CellMarker_enrich_brain_set_res.case_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_EOAD,disease_genes = list_case_earlyAD)

CellMarker_enrich_brain_set_res.PPI_dementia<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_dementia,disease_genes = dementia_PPI_cluster)
CellMarker_enrich_brain_set_res.PPI_AD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_AD,disease_genes = AD_PPI_cluster)
CellMarker_enrich_brain_set_res.PPI_earlyAD<-get_CellMarker_celltypeEnrichment(celltype_markers.CellMarker_set.hgnormal_brain.anno_EOAD,disease_genes = earlyAD_PPI_cluster)
```

# Section 3: filter significant results

### Brain tissues

```{r, message=FALSE, warning=FALSE}

# brain

sig_CellMarker.brain.burdendementia<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.burdendementia,p_adjust_value = 0.1)
sig_CellMarker.brain.burdenAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.burdenAD,p_adjust_value = 0.1)
sig_CellMarker.brain.burdenearlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.burdenearlyAD,p_adjust_value = 0.1)

sig_CellMarker.brain.caseonly_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.caseonly_dementia,p_adjust_value = 0.1)
sig_CellMarker.brain.caseonly_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.caseonly_AD,p_adjust_value = 0.1)
sig_CellMarker.brain.caseonly_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.caseonly_earlyAD,p_adjust_value = 0.1)

sig_CellMarker.brain.case_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.case_dementia,p_adjust_value = 0.1)
sig_CellMarker.brain.case_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.case_AD,p_adjust_value = 0.1)
sig_CellMarker.brain.case_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.case_earlyAD,p_adjust_value = 0.1)

sig_CellMarker.brain.PPI_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.PPI_dementia,p_adjust_value = 0.1)
sig_CellMarker.brain.PPI_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.PPI_AD,p_adjust_value = 0.1)
sig_CellMarker.brain.PPI_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_res.PPI_earlyAD,p_adjust_value = 0.1)

# brain_set

sig_CellMarker.brain_set.burdendementia<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.burdendementia,p_adjust_value = 0.1)
sig_CellMarker.brain_set.burdenAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.burdenAD,p_adjust_value = 0.1)
sig_CellMarker.brain_set.burdenearlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.burdenearlyAD,p_adjust_value = 0.1)

sig_CellMarker.brain_set.caseonly_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.caseonly_dementia,p_adjust_value = 0.1)
sig_CellMarker.brain_set.caseonly_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.caseonly_AD,p_adjust_value = 0.1)
sig_CellMarker.brain_set.caseonly_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.caseonly_earlyAD,p_adjust_value = 0.1)

sig_CellMarker.brain_set.case_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.case_dementia,p_adjust_value = 0.1)
sig_CellMarker.brain_set.case_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.case_AD,p_adjust_value = 0.1)
sig_CellMarker.brain_set.case_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.case_earlyAD,p_adjust_value = 0.1)

sig_CellMarker.brain_set.PPI_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.PPI_dementia,p_adjust_value = 0.1)
sig_CellMarker.brain_set.PPI_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.PPI_AD,p_adjust_value = 0.1)
sig_CellMarker.brain_set.PPI_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_brain_set_res.PPI_earlyAD,p_adjust_value = 0.1)
```

### All tissues

```{r message=FALSE, warning=FALSE}

# all

sig_CellMarker.all.burdendementia<-get_CellMarker_enrichfilter(CellMarker_enrich_res.burdendementia,p_adjust_value = 0.1)
sig_CellMarker.all.burdenAD<-get_CellMarker_enrichfilter(CellMarker_enrich_res.burdenAD,p_adjust_value = 0.1)
sig_CellMarker.all.burdenearlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_res.burdenearlyAD,p_adjust_value = 0.1)

sig_CellMarker.all.caseonly_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_res.caseonly_dementia,p_adjust_value = 0.1)
sig_CellMarker.all.caseonly_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_res.caseonly_AD,p_adjust_value = 0.1)
sig_CellMarker.all.caseonly_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_res.caseonly_earlyAD,p_adjust_value = 0.1)

sig_CellMarker.all.case_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_res.case_dementia,p_adjust_value = 0.1)
sig_CellMarker.all.case_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_res.case_AD,p_adjust_value = 0.1)
sig_CellMarker.all.case_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_res.case_earlyAD,p_adjust_value = 0.1)

sig_CellMarker.all.PPI_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_res.PPI_dementia,p_adjust_value = 0.1)
sig_CellMarker.all.PPI_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_res.PPI_AD,p_adjust_value = 0.1)
sig_CellMarker.all.PPI_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_res.PPI_earlyAD,p_adjust_value = 0.1)

# all_set

sig_CellMarker.all_set.burdendementia<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.burdendementia,p_adjust_value = 0.1)
sig_CellMarker.all_set.burdenAD<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.burdenAD,p_adjust_value = 0.1)
sig_CellMarker.all_set.burdenearlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.burdenearlyAD,p_adjust_value = 0.1)

sig_CellMarker.all_set.caseonly_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.caseonly_dementia,p_adjust_value = 0.1)
sig_CellMarker.all_set.caseonly_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.caseonly_AD,p_adjust_value = 0.1)
sig_CellMarker.all_set.caseonly_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.caseonly_earlyAD,p_adjust_value = 0.1)

sig_CellMarker.all_set.case_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.case_dementia,p_adjust_value = 0.1)
sig_CellMarker.all_set.case_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.case_AD,p_adjust_value = 0.1)
sig_CellMarker.all_set.case_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.case_earlyAD,p_adjust_value = 0.1)

sig_CellMarker.all_set.PPI_dementia<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.PPI_dementia,p_adjust_value = 0.1)
sig_CellMarker.all_set.PPI_AD<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.PPI_AD,p_adjust_value = 0.1)
sig_CellMarker.all_set.PPI_earlyAD<-get_CellMarker_enrichfilter(CellMarker_enrich_set_res.PPI_earlyAD,p_adjust_value = 0.1)
```

## save significant enrichment results

```{r}

# brain res

xl_list_brain<-list()

xl_list_brain$sig_burden_dementia<-sig_CellMarker.brain.burdendementia
xl_list_brain$sig_burden_AD<-sig_CellMarker.brain.burdenAD
xl_list_brain$sig_burden_earlyAD<-sig_CellMarker.brain.burdenearlyAD

xl_list_brain$sig_caseonly_dementia<-sig_CellMarker.brain.caseonly_dementia
xl_list_brain$sig_caseonly_AD<-sig_CellMarker.brain.caseonly_AD
xl_list_brain$sig_caseonly_earlyAD<-sig_CellMarker.brain.caseonly_earlyAD

xl_list_brain$sig_case_dementia<-sig_CellMarker.brain.case_dementia
xl_list_brain$sig_case_AD<-sig_CellMarker.brain.case_AD
xl_list_brain$sig_case_earlyAD<-sig_CellMarker.brain.case_earlyAD

xl_list_brain$sig_PPI_dementia<-sig_CellMarker.brain.PPI_dementia
xl_list_brain$sig_PPI_AD<-sig_CellMarker.brain.PPI_AD
xl_list_brain$sig_PPI_earlyAD<-sig_CellMarker.brain.PPI_earlyAD

# brain_set res

xl_list_brain_set<-list()

xl_list_brain_set$sig_burden_dementia<-sig_CellMarker.brain_set.burdendementia
xl_list_brain_set$sig_burden_AD<-sig_CellMarker.brain_set.burdenAD
xl_list_brain_set$sig_burden_earlyAD<-sig_CellMarker.brain_set.burdenearlyAD

xl_list_brain_set$sig_caseonly_dementia<-sig_CellMarker.brain_set.caseonly_dementia
xl_list_brain_set$sig_caseonly_AD<-sig_CellMarker.brain_set.caseonly_AD
xl_list_brain_set$sig_caseonly_earlyAD<-sig_CellMarker.brain_set.caseonly_earlyAD

xl_list_brain_set$sig_case_dementia<-sig_CellMarker.brain_set.case_dementia
xl_list_brain_set$sig_case_AD<-sig_CellMarker.brain_set.case_AD
xl_list_brain_set$sig_case_earlyAD<-sig_CellMarker.brain_set.case_earlyAD

xl_list_brain_set$sig_PPI_dementia<-sig_CellMarker.brain_set.PPI_dementia
xl_list_brain_set$sig_PPI_AD<-sig_CellMarker.brain_set.PPI_AD
xl_list_brain_set$sig_PPI_earlyAD<-sig_CellMarker.brain_set.PPI_earlyAD

# all res

xl_list_all<-list()

xl_list_all$sig_burden_dementia<-sig_CellMarker.all.burdendementia
xl_list_all$sig_burden_AD<-sig_CellMarker.all.burdenAD
xl_list_all$sig_burden_earlyAD<-sig_CellMarker.all.burdenearlyAD

xl_list_all$sig_caseonly_dementia<-sig_CellMarker.all.caseonly_dementia
xl_list_all$sig_caseonly_AD<-sig_CellMarker.all.caseonly_AD
xl_list_all$sig_caseonly_earlyAD<-sig_CellMarker.all.caseonly_earlyAD

xl_list_all$sig_case_dementia<-sig_CellMarker.all.case_dementia
xl_list_all$sig_case_AD<-sig_CellMarker.all.case_AD
xl_list_all$sig_case_earlyAD<-sig_CellMarker.all.case_earlyAD

xl_list_all$sig_PPI_dementia<-sig_CellMarker.all.PPI_dementia
xl_list_all$sig_PPI_AD<-sig_CellMarker.all.PPI_AD
xl_list_all$sig_PPI_earlyAD<-sig_CellMarker.all.PPI_earlyAD

# all_set res

xl_list_all_set<-list()

xl_list_all_set$sig_burden_dementia<-sig_CellMarker.all_set.burdendementia
xl_list_all_set$sig_burden_AD<-sig_CellMarker.all_set.burdenAD
xl_list_all_set$sig_burden_earlyAD<-sig_CellMarker.all_set.burdenearlyAD

xl_list_all_set$sig_caseonly_dementia<-sig_CellMarker.all_set.caseonly_dementia
xl_list_all_set$sig_caseonly_AD<-sig_CellMarker.all_set.caseonly_AD
xl_list_all_set$sig_caseonly_earlyAD<-sig_CellMarker.all_set.caseonly_earlyAD

xl_list_all_set$sig_case_dementia<-sig_CellMarker.all_set.case_dementia
xl_list_all_set$sig_case_AD<-sig_CellMarker.all_set.case_AD
xl_list_all_set$sig_case_earlyAD<-sig_CellMarker.all_set.case_earlyAD

xl_list_all_set$sig_PPI_dementia<-sig_CellMarker.all_set.PPI_dementia
xl_list_all_set$sig_PPI_AD<-sig_CellMarker.all_set.PPI_AD
xl_list_all_set$sig_PPI_earlyAD<-sig_CellMarker.all_set.PPI_earlyAD

## output result tables

write.xlsx(xl_list_brain,"./results/CellMarker_enrich_sig_brain.xlsx")
write.xlsx(xl_list_brain_set,"./results/CellMarker_enrich_sig_brain_set.xlsx")
write.xlsx(xl_list_all,"./results/CellMarker_enrich_sig_all.xlsx")
write.xlsx(xl_list_all_set,"./results/CellMarker_enrich_sig_all_set.xlsx")
```

################################################################################
#
# plot drawing
#
################################################################################

# prepare data for visualization

## collect all the significant cell types

```{r}

# brain

ls.sig_CellMarker.brain<-unique(c(sig_CellMarker.brain.case_AD$cell_type,sig_CellMarker.brain.case_earlyAD$cell_type,sig_CellMarker.brain.PPI_dementia$cell_type)) %>% sort

# brain_set

ls.sig_CellMarker.brain_set<-unique(c(sig_CellMarker.brain_set.case_AD$cell_type,sig_CellMarker.brain_set.case_earlyAD$cell_type)) %>% sort

# all

ls.sig_CellMarker.all<-unique(c(sig_CellMarker.all.burdendementia$cell_type,sig_CellMarker.all.caseonly_dementia$cell_type,sig_CellMarker.all.caseonly_AD$cell_type,sig_CellMarker.all.case_dementia$cell_type,sig_CellMarker.all.case_AD$cell_type,sig_CellMarker.all.case_earlyAD$cell_type,sig_CellMarker.all.PPI_dementia$cell_type,sig_CellMarker.all.PPI_AD$cell_type,sig_CellMarker.all.PPI_earlyAD$cell_type)) %>% sort

# all_set

ls.sig_CellMarker.all_set<-unique(c(sig_CellMarker.all_set.burdendementia$cell_type,sig_CellMarker.all_set.caseonly_dementia$cell_type,sig_CellMarker.all_set.caseonly_AD$cell_type,sig_CellMarker.all_set.case_dementia$cell_type,sig_CellMarker.all_set.case_AD$cell_type,sig_CellMarker.all_set.case_earlyAD$cell_type,sig_CellMarker.all_set.PPI_dementia$cell_type,sig_CellMarker.all_set.PPI_AD$cell_type,sig_CellMarker.all_set.PPI_earlyAD$cell_type)) %>% sort
```

## add group and combine

### all

```{r message=FALSE, warning=FALSE}
plt.burdendementia<-CellMarker_enrich_res.burdendementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("burdened dementia",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)
plt.burdenAD<-CellMarker_enrich_res.burdenAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("burdened AD",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)
plt.burdenearlyAD<-CellMarker_enrich_res.burdenearlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("burdened EOAD",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)

plt.caseonly_dementia<-CellMarker_enrich_res.caseonly_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("case-only dementia",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)
plt.caseonly_AD<-CellMarker_enrich_res.caseonly_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("case-only AD",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)
plt.caseonly_earlyAD<-CellMarker_enrich_res.caseonly_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("case-only EOAD",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)

plt.case_dementia<-CellMarker_enrich_res.case_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("case dementia",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)
plt.case_AD<-CellMarker_enrich_res.case_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("case AD",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)
plt.case_earlyAD<-CellMarker_enrich_res.case_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("case EOAD",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)

plt.PPI_dementia<-CellMarker_enrich_res.PPI_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("PPI dementia",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)
plt.PPI_AD<-CellMarker_enrich_res.PPI_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("PPI AD",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)
plt.PPI_earlyAD<-CellMarker_enrich_res.PPI_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all) %>% cbind(.,group=rep("PPI EOAD",length(ls.sig_CellMarker.all))) %>% arrange(cell_type)

# res enrich plt combine

plt.all<-rbind(plt.burdendementia,plt.burdenAD,plt.burdenearlyAD,plt.caseonly_dementia,plt.caseonly_AD,plt.caseonly_earlyAD,plt.case_dementia,plt.case_AD,plt.case_earlyAD,plt.PPI_dementia,plt.PPI_AD,plt.PPI_earlyAD)
```

### all_set

```{r message=FALSE, warning=FALSE}
plt.burdendementia<-CellMarker_enrich_set_res.burdendementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("burdened dementia",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)
plt.burdenAD<-CellMarker_enrich_set_res.burdenAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("burdened AD",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)
plt.burdenearlyAD<-CellMarker_enrich_set_res.burdenearlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("burdened EOAD",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)

plt.caseonly_dementia<-CellMarker_enrich_set_res.caseonly_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("case-only dementia",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)
plt.caseonly_AD<-CellMarker_enrich_set_res.caseonly_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("case-only AD",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)
plt.caseonly_earlyAD<-CellMarker_enrich_set_res.caseonly_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("case-only EOAD",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)

plt.case_dementia<-CellMarker_enrich_set_res.case_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("case dementia",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)
plt.case_AD<-CellMarker_enrich_set_res.case_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("case AD",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)
plt.case_earlyAD<-CellMarker_enrich_set_res.case_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("case EOAD",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)

plt.PPI_dementia<-CellMarker_enrich_set_res.PPI_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("PPI dementia",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)
plt.PPI_AD<-CellMarker_enrich_set_res.PPI_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("PPI AD",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)
plt.PPI_earlyAD<-CellMarker_enrich_set_res.PPI_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.all_set) %>% cbind(.,group=rep("PPI EOAD",length(ls.sig_CellMarker.all_set))) %>% arrange(cell_type)

# res enrich plt combine

plt.all_set<-rbind(plt.burdendementia,plt.burdenAD,plt.burdenearlyAD,plt.caseonly_dementia,plt.caseonly_AD,plt.caseonly_earlyAD,plt.case_dementia,plt.case_AD,plt.case_earlyAD,plt.PPI_dementia,plt.PPI_AD,plt.PPI_earlyAD)
```

### brain

```{r message=FALSE, warning=FALSE}
plt.burdendementia<-CellMarker_enrich_brain_res.burdendementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("burdened dementia",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)
plt.burdenAD<-CellMarker_enrich_brain_res.burdenAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("burdened AD",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)
plt.burdenearlyAD<-CellMarker_enrich_brain_res.burdenearlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("burdened EOAD",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)

plt.caseonly_dementia<-CellMarker_enrich_brain_res.caseonly_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("case-only dementia",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)
plt.caseonly_AD<-CellMarker_enrich_brain_res.caseonly_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("case-only AD",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)
plt.caseonly_earlyAD<-CellMarker_enrich_brain_res.caseonly_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("case-only EOAD",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)

plt.case_dementia<-CellMarker_enrich_brain_res.case_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("case dementia",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)
plt.case_AD<-CellMarker_enrich_brain_res.case_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("case AD",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)
plt.case_earlyAD<-CellMarker_enrich_brain_res.case_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("case EOAD",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)

plt.PPI_dementia<-CellMarker_enrich_brain_res.PPI_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("PPI dementia",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)
plt.PPI_AD<-CellMarker_enrich_brain_res.PPI_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("PPI AD",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)
plt.PPI_earlyAD<-CellMarker_enrich_brain_res.PPI_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain) %>% cbind(.,group=rep("PPI EOAD",length(ls.sig_CellMarker.brain))) %>% arrange(cell_type)

# res enrich plt combine

plt.brain<-rbind(plt.burdendementia,plt.burdenAD,plt.burdenearlyAD,plt.caseonly_dementia,plt.caseonly_AD,plt.caseonly_earlyAD,plt.case_dementia,plt.case_AD,plt.case_earlyAD,plt.PPI_dementia,plt.PPI_AD,plt.PPI_earlyAD)
```

### brain_set

```{r message=FALSE, warning=FALSE}
plt.burdendementia<-CellMarker_enrich_brain_set_res.burdendementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("burdened dementia",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)
plt.burdenAD<-CellMarker_enrich_brain_set_res.burdenAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("burdened AD",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)
plt.burdenearlyAD<-CellMarker_enrich_brain_set_res.burdenearlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("burdened EOAD",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)

plt.caseonly_dementia<-CellMarker_enrich_brain_set_res.caseonly_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("case-only dementia",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)
plt.caseonly_AD<-CellMarker_enrich_brain_set_res.caseonly_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("case-only AD",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)
plt.caseonly_earlyAD<-CellMarker_enrich_brain_set_res.caseonly_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("case-only EOAD",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)

plt.case_dementia<-CellMarker_enrich_brain_set_res.case_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("case dementia",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)
plt.case_AD<-CellMarker_enrich_brain_set_res.case_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("case AD",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)
plt.case_earlyAD<-CellMarker_enrich_brain_set_res.case_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("case EOAD",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)

plt.PPI_dementia<-CellMarker_enrich_brain_set_res.PPI_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("PPI dementia",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)
plt.PPI_AD<-CellMarker_enrich_brain_set_res.PPI_AD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("PPI AD",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)
plt.PPI_earlyAD<-CellMarker_enrich_brain_set_res.PPI_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_CellMarker.brain_set) %>% cbind(.,group=rep("PPI EOAD",length(ls.sig_CellMarker.brain_set))) %>% arrange(cell_type)

# res enrich plt combine

plt.brain_set<-rbind(plt.burdendementia,plt.burdenAD,plt.burdenearlyAD,plt.caseonly_dementia,plt.caseonly_AD,plt.caseonly_earlyAD,plt.case_dementia,plt.case_AD,plt.case_earlyAD,plt.PPI_dementia,plt.PPI_AD,plt.PPI_earlyAD)
```

## output significant result tables

```{r}
xl_list_sigvisual<-list()

xl_list_sigvisual$all<-plt.all
xl_list_sigvisual$all_set<-plt.all_set
xl_list_sigvisual$brain<-plt.brain
xl_list_sigvisual$brain_set<-plt.brain_set

write.xlsx(xl_list_sigvisual,"./results/CellMarker_sigvisual_table_allgroups.xlsx")
```

# data visualization

## plot drawing

```{r, message=FALSE, warning=FALSE}
heatmap.plt.all<-get_heatmatrix(celltype_enrich=plt.all,ls.sig_CellMarker=ls.sig_CellMarker.all,testing_set_name="all tissues")

heatmap.plt.all_set<-get_heatmatrix(celltype_enrich=plt.all_set,ls.sig_CellMarker=ls.sig_CellMarker.all_set,testing_set_name="all tissues from set")

heatmap.plt.brain<-get_heatmatrix(celltype_enrich=plt.brain,ls.sig_CellMarker=ls.sig_CellMarker.brain,testing_set_name="brain tissues")

heatmap.plt.brain_set<-get_heatmatrix(celltype_enrich=plt.brain_set,ls.sig_CellMarker=ls.sig_CellMarker.brain_set,testing_set_name="brain tissues from set")
```

## output the plots

```{r}
# define trellis object saving function

save_trellis <- function(heatmap_plot,ls.sig_CellMarker,file_name,save_path){
 pdf(paste(save_path,file_name,sep = '/'),height=20,width=2+20*length(ls.sig_CellMarker)/12)
 print(heatmap_plot)
 dev.off()
}

# save plot
save_trellis(heatmap.plt.all,ls.sig_CellMarker.all,"Plots/plot_single-cell_enrichment_CellMarker-all.pdf",outPath)
save_trellis(heatmap.plt.all_set,ls.sig_CellMarker.all_set,"Plots/plot_single-cell_enrichment_CellMarker-all_set.pdf",outPath)
save_trellis(heatmap.plt.brain,ls.sig_CellMarker.brain,"Plots/plot_single-cell_enrichment_CellMarker-brain.pdf",outPath)
save_trellis(heatmap.plt.brain_set,ls.sig_CellMarker.brain_set,"Plots/plot_single-cell_enrichment_CellMarker-brain_set.pdf",outPath)
```


## output the overlapped genes in the enriched brain cell types 

### filter significant brain markers overlapped genes

```{r}
sig_brain_markers.anno_AD<-celltype_markers.CellMarker.hgnormal_brain.anno_AD %>% dplyr::select(entrezgene_id,GENE,all_of(ls.sig_CellMarker.brain)) %>% dplyr::filter((get(ls.sig_CellMarker.brain[1])==1)|(get(ls.sig_CellMarker.brain[2])==1))
sig_brain_set_markers.anno_AD<-celltype_markers.CellMarker_set.hgnormal_brain.anno_AD %>% dplyr::select(entrezgene_id,GENE,all_of(ls.sig_CellMarker.brain_set)) %>% dplyr::filter((get(ls.sig_CellMarker.brain_set[1])==1)|(get(ls.sig_CellMarker.brain_set[2])==1))

sig_brain_markers.anno_EOAD<-celltype_markers.CellMarker.hgnormal_brain.anno_EOAD %>% dplyr::select(entrezgene_id,GENE,all_of(ls.sig_CellMarker.brain)) %>% dplyr::filter((get(ls.sig_CellMarker.brain[1])==1)|(get(ls.sig_CellMarker.brain[2])==1))
sig_brain_set_markers.anno_EOAD<-celltype_markers.CellMarker_set.hgnormal_brain.anno_EOAD %>% dplyr::select(entrezgene_id,GENE,all_of(ls.sig_CellMarker.brain_set)) %>% dplyr::filter((get(ls.sig_CellMarker.brain_set[1])==1)|(get(ls.sig_CellMarker.brain_set[2])==1))

sig_brain_markers.anno_dementia<-celltype_markers.CellMarker.hgnormal_brain.anno_dementia %>% dplyr::select(entrezgene_id,GENE,all_of(ls.sig_CellMarker.brain)) %>% dplyr::filter((get(ls.sig_CellMarker.brain[1])==1)|(get(ls.sig_CellMarker.brain[2])==1))
```

### filter specific gene sets overlapped significant brain markers

```{r}
# method 1

overlap.sig_brain_markers_caseAD<-attr(venn(list(sig_brain_markers.anno_AD$entrezgene_id,list_case_AD)), "intersections")
overlap.sig_brain_markers_caseEOAD<-attr(venn(list(sig_brain_markers.anno_EOAD$entrezgene_id,list_case_earlyAD)), "intersections")
overlap.sig_brain_set_markers_caseAD<-attr(venn(list(sig_brain_set_markers.anno_AD$entrezgene_id,list_case_AD)), "intersections")
overlap.sig_brain_set_markers_caseEOAD<-attr(venn(list(sig_brain_set_markers.anno_EOAD$entrezgene_id,list_case_earlyAD)), "intersections")

dementia_PPI_cluster<-read.table("BIOGRID_PPI/dementia_PPI_cluster.txt")[,1]
overlap.sig_brain_markers_PPIdementia<-attr(venn(list(sig_brain_markers.anno_dementia$entrezgene_id,dementia_PPI_cluster)), "intersections")

overlap.sig_brain_markers_caseAD<-overlap.sig_brain_markers_caseAD$`A:B`
overlap.sig_brain_markers_caseEOAD<-overlap.sig_brain_markers_caseEOAD$`A:B`
overlap.sig_brain_set_markers_caseAD<-overlap.sig_brain_set_markers_caseAD$`A:B`
overlap.sig_brain_set_markers_caseEOAD<-overlap.sig_brain_set_markers_caseEOAD$`A:B`
overlap.sig_brain_markers_PPIdementia<-overlap.sig_brain_markers_PPIdementia$`A:B`

# method 2

overlap.sig_brain_markers_caseAD<-sig_brain_markers.anno_AD %>% dplyr::filter(entrezgene_id %in% list_case_AD)
overlap.sig_brain_markers_caseEOAD<-sig_brain_markers.anno_EOAD %>% dplyr::filter(entrezgene_id %in% list_case_earlyAD)
overlap.sig_brain_set_markers_caseAD<-sig_brain_set_markers.anno_AD %>% dplyr::filter(entrezgene_id %in% list_case_AD)
overlap.sig_brain_set_markers_caseEOAD<-sig_brain_set_markers.anno_EOAD %>% dplyr::filter(entrezgene_id %in% list_case_earlyAD)

overlap.sig_brain_markers_PPIdementia<-sig_brain_markers.anno_dementia %>% dplyr::filter(entrezgene_id %in% dementia_PPI_cluster)
```

### output results

```{r}
overlap.sig_brain.list<-list()
overlap.sig_brain.list$sigbrain_caseAD<-overlap.sig_brain_markers_caseAD
overlap.sig_brain.list$sigbrain_caseEOAD<-overlap.sig_brain_markers_caseEOAD
overlap.sig_brain.list$sigbrain_set_caseAD<-overlap.sig_brain_set_markers_caseAD
overlap.sig_brain.list$sigbrain_set_caseEOAD<-overlap.sig_brain_set_markers_caseEOAD
overlap.sig_brain.list$sigbrain_PPIdementia<-overlap.sig_brain_markers_PPIdementia
write.xlsx(overlap.sig_brain.list,paste(outPath,"case_genesets_overlapped_significant_brainmarkers.xlsx",sep = '/'))
```

## add evidence to the ranking list

```{r}
# the previous script in "WGCNA_brainregioncombined_enrichment_analysis-GO"

evidence_df<-read.xlsx("./metadata/evidence-ranking_table_all-case-only-and-burdened-genes.xlsx")

## overlap with brain cell type markers or not

### brain markers
gene_list<-celltype_markers.CellMarker.hgnormal_brain$geneID %>% unique # 600
evidence_df$CellMarker_brainmarker_overlap<-evidence_df$ENTREZID %>% lapply(.,function(x){if(x %in% gene_list){1}else{0}}) %>% unlist

### brain markers from set
gene_list<-celltype_markers.CellMarker_set.hgnormal_brain$geneID %>% unique # 592
evidence_df$CellMarker_set_brainmarker_overlap<-evidence_df$ENTREZID %>% lapply(.,function(x){if(x %in% gene_list){1}else{0}}) %>% unlist

## enriched in the brain-related cell-types or not

write.xlsx(evidence_df,"./metadata/evidence-ranking_table_all-case-only-and-burdened-genes.xlsx")

# the next script is "ALLENexpdata_exploration"
```

## save.image

```{r}
save.image("./singlecell_marker_enrichment_analysis_CellMarker.RData")
```
