---
title: "spatiotemporal enrichment: data preprocessing"
output: html_notebook
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## load packages & data

```{r, message=FALSE, warning=FALSE}
library(openxlsx)
library(scatterplot3d)
library(biomaRt)
library(stringr)
library(dplyr)
library(stats)
library(reshape2)
library(forcats)
library(magrittr)

options(stringsAsFactors = FALSE)
source("function_spatiotemporal_enrichment_plots.R")
load("burdengenes_PPIpartner_extraction.RData")
```

#===============================================================================
#
#  Code chunk 1: spatio-temporal data loading and preprocessing
#
#===============================================================================

## spatio-temporal data input: brainspan

```{r}
Brainspan_RNAseq <- read.csv("./data/Brainspan/genes_matrix_csv/expression_matrix.csv",header = F)
Brainspan_RNAseq <- Brainspan_RNAseq[,-c(1)]
Brainspan_RNAseq_col <- read.csv("./data/Brainspan/genes_matrix_csv/columns_metadata.csv")
Brainspan_RNAseq_row <- read.csv("./data/Brainspan/genes_matrix_csv/rows_metadata.csv")
```

## data preprocessing: annotate development periods

```{r}
Brainspan_RNAseq_col.age <- Brainspan_RNAseq_col$age
Brainspan_RNAseq_col.age <- as.data.frame(matrix(unlist(strsplit(Brainspan_RNAseq_col$age," ")),ncol=2,byrow=TRUE,dimnames = list(c(1:length(Brainspan_RNAseq_col$age)),c("number", "unit"))))
Brainspan_RNAseq_col.age$number <- as.integer(Brainspan_RNAseq_col.age$number)
Brainspan_RNAseq_col$developmental_period<-get_temperalperiod_list(Brainspan_RNAseq_col.age)

Brainspan_RNAseq_col.age$structure <- Brainspan_RNAseq_col$structure_acronym
Brainspan_RNAseq_col.age$developmental_period <- Brainspan_RNAseq_col$developmental_period
Brainspan_RNAseq_col.age$structure_name <- Brainspan_RNAseq_col$structure_name
table(Brainspan_RNAseq_col.age$developmental_period)
```

## data inspection

### spatio-temporal distribution of available samples

```{r}
period_stucture_table<-get_period_stucture(Brainspan_RNAseq_col.age)

write.xlsx(Brainspan_RNAseq_col.age,"./results/Brainspan_RNAseq_col.age_structure.xlsx")
write.xlsx(period_stucture_table,"./results/Brainspan_RNAseq_period_stucture.xlsx",row.names=TRUE)
```

### prepare expression matrix using period 3-7 for hierarchical clustering of brain regions

```{r}
# period 3-7 data extraction for clustering (brain-region devided)

period3_7_regionexpression_table <- get_hierarchical_expression(Brainspan_RNAseq_col.age,Brainspan_RNAseq_row,Brainspan_RNAseq)

# period 3-7 data preprocessing for regional hierarchical clustering

period3_7_regionexpression_table <- period3_7_regionexpression_table[,order(period3_7_regionexpression_table[1,])]
period3_7_brainregion_table<-as.data.frame(table(as.character(period3_7_regionexpression_table[1,])))
region_name<-unique(as.character(period3_7_regionexpression_table[1,]))
period3_7_brainregion_table.col<-c()
for(i in 1:length(region_name)){
  number_sample<-paste(region_name[i],seq(from=1, to=period3_7_brainregion_table[which(period3_7_brainregion_table$Var1==region_name[i]),"Freq"], by=1),sep = "_")
  period3_7_brainregion_table.col<-c(period3_7_brainregion_table.col,number_sample)
}

colnames(period3_7_regionexpression_table)<- period3_7_brainregion_table.col
period3_7_regionexpression_table<-period3_7_regionexpression_table[-c(1),]
period3_7_regionexpression_table<-apply(period3_7_regionexpression_table,1,as.numeric)
period3_7_regionexpression_table<-as.data.frame(t(period3_7_regionexpression_table))
colnames(period3_7_regionexpression_table)<- period3_7_brainregion_table.col
x=as.numeric(apply(period3_7_regionexpression_table,1,sum)); x<-as.data.frame(x)
period3_7_regionexpression_table<-cbind(period3_7_regionexpression_table,x)
colnames(period3_7_regionexpression_table)<- c(period3_7_brainregion_table.col,"sum_expression")
period3_7_regionexpression_table<-period3_7_regionexpression_table[which(period3_7_regionexpression_table$sum_expression!=0),]
x=as.numeric(rownames(period3_7_regionexpression_table[which(period3_7_regionexpression_table$sum_expression<=0),]))
summary(as.numeric(period3_7_regionexpression_table$sum_expression))

period3_7_regionexpression_table<-period3_7_regionexpression_table[,1:ncol(period3_7_regionexpression_table)-1]

# output result table

write.csv(period3_7_regionexpression_table,"./results/period3_7_regionexpression_table.csv",row.names = F)
```

#===============================================================================
#
#  Code chunk 2: PCA de-outlier
#
#===============================================================================

## PCA test for samples from period 3-7

```{r}
probesetvar = apply(period3_7_regionexpression_table, 1, var)
ord = order(probesetvar, decreasing=TRUE)
pca = prcomp(t(period3_7_regionexpression_table[ord,]), scale=TRUE)
ss=summary(pca)
pca_table<-as.data.frame(pca$x);pca_table<-apply(pca_table,2,as.numeric)
rownames(pca_table)<-colnames(period3_7_regionexpression_table)

write.xlsx(pca_table,"./results/Spatiotemperal_RNAseq_period37_PCAanalysis.xlsx",row.names=T)
```

## PCA plot drawing for samples from period 3-7

```{r}
var_explained <- pca$sdev^2/sum(pca$sdev^2)

par(mfrow = c( 1, 1 ))
plot(pca$x[,1:2],type="p",xlab=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
     ylab=paste0("PC2: ",round(var_explained[2]*100,1),"%"),
     col=4,pch=21,main="sample PCA figure",cex.main=1)
text(pca$x,labels = rownames(pca$x),pos = 1,offset = 0.6,cex = 0.3)

scatterplot3d(pca$x[,1:3],pch=c(20),color = 4)
text(pca$x[,1:3],labels = rownames(pca$x),cex = 0.5,pos=c(1),col="black")
text(pca$x,labels = rownames(pca$x),pos = 4,offset = 0.6,cex = 0.5)
```

## hierarchical clustering according to regional similarity and melting the data

### merge PCs of samples from the same brain regions via median values

```{r}
pca_table_merge<-pca_table[,c(1:198)]
pca_table_merge<-t(pca_table_merge)
brainregion_clustertable<-get_brainregion_clustertable(period3_7_regionexpression_table)
write.xlsx(brainregion_clustertable,"./results/brainregion_period3_7_clustertable.xlsx",row.names=T)
```

### apply hclust to different brain regions

```{r}
dd <- as.dist(1 - cor(t(brainregion_clustertable),method ="spearman")^2)
hc<-hclust(dd,method = "ward.D")
plot(hc)
```

## process the data according to the hierarchical clustering results

```{r}
# denote the cluster components according to the clustering results

cluster1<-c("A1C","ITC","STC","V1C")
cluster2<-c("DFC","MFC","OFC","VFC")
cluster3<-c("M1C","S1C","IPC")
cluster4<-c("MD","CBC")
cluster5<-c("STR","AMY","HIP")
cluster_list<-list(cluster1,cluster2,cluster3,cluster4,cluster5)

# delete samples from the ten brain regions that have null value in more than half of the periods

regionselected_list<-c("Ocx","M1C-S1C","MGE","URL","CGE","DTH","LGE","PCx","TCx","CB")
Brainspan_RNAseq<-deregion_Brainspan_RNAseq(regionselected_list,Brainspan_RNAseq,Brainspan_RNAseq_col)
Brainspan_RNAseq_col<-deregion_Brainspan_RNAseq.col(regionselected_list,Brainspan_RNAseq_col)
Brainspan_RNAseq_col<-mark_cluster_expressiontable(cluster_list,Brainspan_RNAseq_col)

# test repeated region samples from the same donors

test_table<-Brainspan_RNAseq_col[,c("donor_id","structure_acronym")]
test_table<-test_table[order(test_table$donor_id),]
test_table<-paste(test_table$donor_id,test_table$structure_acronym,sep = "_")
table(test_table)
## no repeated samples exist
rm(test_table)

# remove genes with zero expression value across all remaining brain regions

x=as.numeric(apply(Brainspan_RNAseq,1,sum))
x<-as.data.frame(x)
Brainspan_RNAseq<-cbind(Brainspan_RNAseq,x)
rm_rownames<-rownames(Brainspan_RNAseq[which(Brainspan_RNAseq$x==0),])
Brainspan_RNAseq_row<-Brainspan_RNAseq_row[!(rownames(Brainspan_RNAseq_row) %in% rm_rownames),]
Brainspan_RNAseq<-Brainspan_RNAseq[!(rownames(Brainspan_RNAseq) %in% rm_rownames),]
Brainspan_RNAseq<-Brainspan_RNAseq[,1:length(Brainspan_RNAseq[1,])-1]
rm(rm_rownames)
rownames(Brainspan_RNAseq_row)<-c(1:length(Brainspan_RNAseq_row[,1]))
Brainspan_RNAseq_row$row_num<-c(1:length(Brainspan_RNAseq_row[,1]))
Brainspan_RNAseq_col$column_num<-c(1:length(Brainspan_RNAseq_col[,1]))
```

## PCA test for all remaining samples

```{r}
probesetvar = apply(Brainspan_RNAseq, 1, var)
ord = order(probesetvar, decreasing=TRUE)
pca = prcomp(t(Brainspan_RNAseq[ord,]), scale=TRUE)
ss=summary(pca)
pca_table<-as.data.frame(pca$x)

var_explained <- pca$sdev^2/sum(pca$sdev^2)
var_explained[1:3]
sum(var_explained[1:37])
# 0.6019478

par(mfrow = c( 1, 1 ))
plot(pca$x[,1:2],type="p",xlab=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
     ylab=paste0("PC2: ",round(var_explained[2]*100,1),"%"),
     col=4,pch=21,main="sample PCA figure",cex.main=1)
text(pca$x,labels = rownames(pca$x),pos = 2,offset = 0.6,cex = 0.3)

scatterplot3d(pca$x[,1:3],pch=c(20),color = 4)
text(pca$x[,1:3],labels = rownames(pca$x),cex = 0.5,pos=c(1),col="black")
text(pca$x,labels = rownames(pca$x),pos = 4,offset = 0.6,cex = 0.5)
pca_table<-pca$x;rownames(pca_table)<-c(1:length(pca$x[,1]))
pca_table<-cbind(pca_table,as.data.frame(Brainspan_RNAseq_col$structure_acronym))
write.csv(pca_table,"./results/Brainspan_RNAseq_PCA.csv")
```

## PCA outlier detection & removal

### outlier detection: define outliers using six-time standard deviations

```{r}
var_explained[1:10]
sum(var_explained[1:7])
outliers=apply(pca_table[,1:7],2,function(x)which(abs(x-mean(x))>6*sd(x)))
outliers

save(list = c("pca_table","pca","var_explained"),file = "./spatiotemporal_enrichment_pcavalues.RData")
```

### outlier removal: DFC,MFC

```{r}
Brainspan_RNAseq<-Brainspan_RNAseq[,-c(331,339)]
Brainspan_RNAseq_col<-Brainspan_RNAseq_col[-c(331,339),]
Brainspan_RNAseq_col$column_num<-c(1:length(Brainspan_RNAseq_col[,1]))

# remove genes with zero expression value across all remaining brain regions

x=as.numeric(apply(Brainspan_RNAseq,1,sum))
x<-as.data.frame(x)
Brainspan_RNAseq<-cbind(Brainspan_RNAseq,x)
rm_rownames<-rownames(Brainspan_RNAseq[which(Brainspan_RNAseq$x==0),])
Brainspan_RNAseq_row<-Brainspan_RNAseq_row[!(rownames(Brainspan_RNAseq_row) %in% rm_rownames),]
Brainspan_RNAseq<-Brainspan_RNAseq[!(rownames(Brainspan_RNAseq) %in% rm_rownames),]
Brainspan_RNAseq<-Brainspan_RNAseq[,1:length(Brainspan_RNAseq[1,])-1]
rm(rm_rownames)
rownames(Brainspan_RNAseq_row)<-c(1:length(Brainspan_RNAseq_row[,1]))
Brainspan_RNAseq_row$row_num<-c(1:length(Brainspan_RNAseq_row[,1]))
Brainspan_RNAseq_col$column_num<-c(1:length(Brainspan_RNAseq_col[,1]))

write.csv(Brainspan_RNAseq,"./results/Brainspan_RNAseq_for_analysis.csv")
write.csv(Brainspan_RNAseq_col,"./results/Brainspan_RNAseq_col_for_analysis.csv",row.names = F)
write.csv(Brainspan_RNAseq_row,"./results/Brainspan_RNAseq_row_for_analysis.csv",row.names = F)
```

## annotate BrainSpan genes without entriz id using metascape

## after annotation, delete genes without entriz id

```{r}
Brainspan_RNAseq_row<-read.csv("./results/Brainspan_RNAseq_row_for_analysis.csv")

x=Brainspan_RNAseq_row[which(Brainspan_RNAseq_row$entrez_id!="#N/A"),"row_num"]
Brainspan_RNAseq<-Brainspan_RNAseq[x,]
Brainspan_RNAseq_row<-Brainspan_RNAseq_row[x,]
Brainspan_RNAseq_row$row_num<-c(1:length(Brainspan_RNAseq_row$row_num))
x=Brainspan_RNAseq_row[which(Brainspan_RNAseq_row$entrez_id!="0"),"row_num"]
Brainspan_RNAseq<-Brainspan_RNAseq[x,]
Brainspan_RNAseq_row<-Brainspan_RNAseq_row[x,]
Brainspan_RNAseq_row$row_num<-c(1:length(Brainspan_RNAseq_row$row_num))
colnames(Brainspan_RNAseq)<-paste("num",seq(from=1,to = length(Brainspan_RNAseq[1,]),by = 1),sep = "_")
rownames(Brainspan_RNAseq)<-Brainspan_RNAseq_row$ensembl_gene_id

mart <- useDataset ("hsapiens_gene_ensembl", useMart ("ensembl")) 
genes <- getBM(attributes=c ("ensembl_gene_id", "entrezgene_id"),filters="ensembl_gene_id", values=Brainspan_RNAseq_row$ensembl_gene_id, mart=mart) 
rm(mart)

write.csv(Brainspan_RNAseq,"./results/Brainspan_RNAseq_for_analysis_denoidgenes.csv")
write.csv(Brainspan_RNAseq_row,"./results/Brainspan_RNAseq_row_for_analysis_denoidgenes.csv",row.names = F)
write.csv(Brainspan_RNAseq_col,"./results/Brainspan_RNAseq_col_for_analysis_denoidgenes.csv",row.names = F)
```

### combine the multiple ensembl gene ids mapping to the same entriz id via selecting the representative one with highest expression value

```{r}
retain_ensembl<-getuniqueentrez(Brainspan_RNAseq_row,Brainspan_RNAseq)
Brainspan_RNAseq$entrez_id<-Brainspan_RNAseq_row$entrez_id
Brainspan_RNAseq<-Brainspan_RNAseq[which(rownames(Brainspan_RNAseq) %in% retain_ensembl),]
Brainspan_RNAseq_row<-Brainspan_RNAseq_row[which(Brainspan_RNAseq_row$ensembl_gene_id %in% retain_ensembl),]
rownames(Brainspan_RNAseq)<-Brainspan_RNAseq$entrez_id
Brainspan_RNAseq<-Brainspan_RNAseq[,1:ncol(Brainspan_RNAseq)-1]
```

#===============================================================================
#
#  Code chunk 3: subset extraction for co-expression network construction
#
#===============================================================================

## optional: inspect the overlap between testing gene sets and Brainspan RNAseq expressed genes

```{r}
table(burdengenes_AD_list$ENTREZID %in% Brainspan_RNAseq_row$entrez_id)
table(burdengenes_ADearly_list$ENTREZID %in% Brainspan_RNAseq_row$entrez_id)
table(burdengenes_dementia_list$ENTREZID %in% Brainspan_RNAseq_row$entrez_id)
table(AD_PPI_cluster %in% Brainspan_RNAseq_row$entrez_id)
table(dementia_PPI_cluster %in% Brainspan_RNAseq_row$entrez_id)
table(ADearly_PPI_cluster %in% Brainspan_RNAseq_row$entrez_id)
```

## Step 1: devide the expression matrix according to the sliding windows and regional clusters

```{r}
window_len=3
sum_windows=13-(window_len-1)
temperal_list<-unique(Brainspan_RNAseq_col$developmental_period)
cluster_list<-unique(Brainspan_RNAseq_col$cluster)

table_list=c()
for(i in 1:length(cluster_list)){
  for(j in 1:sum_windows){
    table_list<-c(table_list,paste("table_cluster",i,"_win",j,sep=""))
  }
}

for(i in 1:length(cluster_list)){
  for(j in 1:sum_windows){
    table_cluster<-as.data.frame(Brainspan_RNAseq_row$entrez_id)
    for(k in 1:length(Brainspan_RNAseq_col$donor_id)){
      if((Brainspan_RNAseq_col[k,"cluster"]==cluster_list[i]) & (Brainspan_RNAseq_col[k,"developmental_period"] %in% temperal_list[j:(j+window_len-1)])){
        table_cluster<-cbind(table_cluster,Brainspan_RNAseq[,k])
      }
    }
    table_cluster<-table_cluster[,-c(1)]
    rownames(table_cluster)<-rownames(Brainspan_RNAseq)
    #de zero expression
    x=as.numeric(apply(table_cluster,1,sum))
    x<-as.data.frame(x)
    table_cluster<-cbind(table_cluster,x)
    table_cluster<-table_cluster[which(table_cluster$x!=0),]
    table_cluster<-table_cluster[,1:length(table_cluster[1,])-1]
    assign(paste("table_cluster",i,"_win",j,sep=""),table_cluster)
  }
}
```

## Step 2: inspect the variance distribution for gene filtering

```{r}
tmp.variance<-c()
for(i in 1:5){
  for(j in 1:sum_windows){
    table_cluster<-get(paste("table_cluster",i,"_win",j,sep=""))
    tmp<-screen_table_cluster_var(table_cluster)$variance
    tmp.variance<-c(tmp.variance,tmp)
  }
}
D<-sort(tmp.variance)
summary(D)
quantile(D,c(0.10,0.01,0.001))
# thus, select the 0.001 quantile for downstream filtering: variance > 0.001 (1e-03)
```

## Step 3: check sample sizes (if smaller than 3, the subsets should be removed from further analyses in the next scripts)

```{r}
check_sample_res<-c()
for(i in 1:length(cluster_list)){
  for(j in 1:sum_windows){
    table_cluster<-get(paste("table_cluster",i,"_win",j,sep=""))
    tmp_res<-list(c(length(table_cluster[1,]),paste("table_cluster",i,"_win",j,sep="")))
    check_sample_res<-c(check_sample_res,tmp_res)
  }
}
check_sample_res<-t(as.data.frame(check_sample_res))
colnames(check_sample_res)<-c("sample_size","spatiotemporal")
check_sample_res<-as.data.frame(check_sample_res)
check_sample_res$sample_size<-as.numeric(check_sample_res$sample_size)

# report the subset to be removed

check_sample_res[which(check_sample_res$sample_size<=3),"spatiotemporal"]
## only "table_cluster5_win1"
```

## optional: check data distribution

```{r}
# Shapiro-Wilk normality test
shapiro.test(as.numeric(table_cluster5_win3[2,]))
shapiro.test(as.numeric(table_cluster5_win2[2,]))
shapiro.test(as.numeric(table_cluster5_win10[2,]))
shapiro.test(as.numeric(table_cluster1_win2[2,]))
hist(sort(as.numeric(table_cluster1_win2[2,])),freq = TRUE)

# conclusion: most do not obey normal distribution, thus need spearman test for correlation analyses
```

## Step 4: screen data via variance threshold and prepare subsets for further analyses

```{r}
for(i in 1:length(cluster_list)){
  for(j in 1:sum_windows){
    table_cluster<-get(paste("table_cluster",i,"_win",j,sep=""))
    table_cluster<-screen_table_cluster_var(table_cluster)
    table_cluster<-table_cluster[which(table_cluster$variance>0.001),]
    assign(paste("table_cluster",i,"_win",j,sep=""),table_cluster)
  }
}

for(i in 1:length(cluster_list)){
  for(j in 1:sum_windows){
    table_cluster<-get(paste("table_cluster",i,"_win",j,sep=""))
    table_cluster<-subset(table_cluster, select=-c(variance))
    assign(paste("table_cluster",i,"_win",j,sep=""),table_cluster)
  }
}
```

################################################################################
#
# Next scripts: scripts of co-expression network building and permutation test
#
################################################################################

## save image

```{r}
save.image("spatiotemporal_enrichment_plots_varscreen.RData")
```
