---
title: "WGCNA analysis using RNA-seq data from Allen Human Brain Atlas"
output: html_notebook
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## load packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(WGCNA)
library(openxlsx)
library(magrittr)
library(operators)
library(lattice)
library(RColorBrewer)
library(clusterProfiler)
library(org.Hs.eg.db)
library(stringr)
library(gplots)

source("./R/function_WGCNA_brainregioncombined_enrichment_analysis-GO.R")
outPath <- "./results"
```

################################################################################
#
# module-trait association result exploration
#
################################################################################

## load data

```{r}
load("./Consensus-Network_module-trait-association.RData",sep = '/'))
setLabels = c("TCx","HIP","PCx") # corresponding to the three sub data frames in order
```

## obtain significant trait-associated modules

### single-trait specific

```{r}
modulespecific.TCx.Disease_state<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 1,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c(">=","<",">=",">="))
modulespecific.HIP.Disease_state<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 2,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c(">=","<",">=",">="))
modulespecific.PCx.Disease_state<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 3,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c(">=","<",">=",">=")) # not exist!

modulespecific.TCx.age<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 1,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("<",">=",">=",">="))
modulespecific.HIP.age<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 2,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("<",">=",">=",">="))
modulespecific.PCx.age<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 3,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("<",">=",">=",">="))

modulespecific.TCx.apo_e4<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 1,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c(">=",">=",">=","<"))
modulespecific.HIP.apo_e4<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 2,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c(">=",">=",">=","<")) # not exist!
modulespecific.PCx.apo_e4<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 3,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c(">=",">=",">=","<"))

Disease_state<-rbind(modulespecific.TCx.Disease_state,modulespecific.HIP.Disease_state,modulespecific.PCx.Disease_state) %>% dplyr::select("MEColorNames") %>% table %>% as.data.frame
Disease_state$trait<-rep("Disease_state",nrow(Disease_state))
age<-rbind(modulespecific.TCx.age,modulespecific.HIP.age,modulespecific.PCx.age) %>% dplyr::select("MEColorNames") %>% table %>% as.data.frame
age$trait<-rep("age",nrow(age))
apo_e4<-rbind(modulespecific.TCx.apo_e4,modulespecific.HIP.apo_e4,modulespecific.PCx.apo_e4) %>% dplyr::select("MEColorNames") %>% table %>% as.data.frame
apo_e4$trait<-rep("apo_e4_allele",nrow(apo_e4))

modulespecific.traits<-rbind(Disease_state,age,apo_e4) %>% dplyr::rename(c(MEColorNames=.)) %>% dplyr::filter(MEColorNames!="not exist!") %>% arrange(desc(Freq))
rm(Disease_state,age,apo_e4)
unique(modulespecific.traits$MEColorNames)
```

### age or disease-state associated modules

```{r}
age_disease.TCx<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 1,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("","<",">=",">="))
age_disease.HIP<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 2,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("","<",">=",">="))
age_disease.PCx<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 3,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("","<",">=",">="))

age_disease.TCx %<>% rbind(.,get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 1,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("<","",">=",">=")))
age_disease.HIP %<>% rbind(.,get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 2,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("<","",">=",">=")))
age_disease.PCx %<>% rbind(.,get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 3,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("<","",">=",">=")))

age_disease.traits<-rbind(age_disease.TCx,age_disease.HIP,age_disease.PCx) %>% dplyr::select("MEColorNames") %>% table %>% as.data.frame %>% dplyr::rename(c(MEColorNames=.)) %>% dplyr::filter(MEColorNames!="not exist!") %>% arrange(desc(Freq))
age_disease.traits$trait<-rep("age or disease-state",nrow(age_disease.traits))

unique(age_disease.traits$MEColorNames)
```

### disease-state associated modules

```{r}
disease.TCx<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 1,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("","<",">=",">="))
disease.HIP<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 2,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("","<",">=",">="))
disease.PCx<-get_sig_trait_Module(moduleTraitPvalue,moduleTraitCor,set_num = 3,trait = c("age","Disease_state","sex","apo_e4_allele"),cutoff_pval = 0.1,largerorsmaller = c("","<",">=",">=")) # "not exist!"

disease.traits<-rbind(disease.TCx,disease.HIP,disease.PCx) %>% dplyr::select("MEColorNames") %>% table %>% as.data.frame %>% dplyr::rename(c(MEColorNames=.)) %>% dplyr::filter(MEColorNames!="not exist!") %>% arrange(desc(Freq))
disease.traits$trait<-rep("disease-state",nrow(disease.traits))

unique(disease.traits$MEColorNames)
```

### output module results

```{r}
modulespecific.TCx.Disease_state$brain<-rep("TCx",nrow(modulespecific.TCx.Disease_state))
modulespecific.HIP.Disease_state$brain<-rep("HIP",nrow(modulespecific.HIP.Disease_state))
modulespecific.TCx.Disease_state$trait<-rep("Disease_state",nrow(modulespecific.TCx.Disease_state))
modulespecific.HIP.Disease_state$trait<-rep("Disease_state",nrow(modulespecific.HIP.Disease_state))
modulespecific.TCx.age$brain<-rep("TCx",nrow(modulespecific.TCx.age))
modulespecific.HIP.age$brain<-rep("HIP",nrow(modulespecific.HIP.age))
modulespecific.PCx.age$brain<-rep("PCx",nrow(modulespecific.PCx.age))
modulespecific.TCx.age$trait<-rep("age",nrow(modulespecific.TCx.age))
modulespecific.HIP.age$trait<-rep("age",nrow(modulespecific.HIP.age))
modulespecific.PCx.age$trait<-rep("age",nrow(modulespecific.PCx.age))
modulespecific.TCx.apo_e4$brain<-rep("TCx",nrow(modulespecific.TCx.apo_e4))
modulespecific.PCx.apo_e4$brain<-rep("PCx",nrow(modulespecific.PCx.apo_e4))
modulespecific.TCx.apo_e4$trait<-rep("apo_e4",nrow(modulespecific.TCx.apo_e4))
modulespecific.PCx.apo_e4$trait<-rep("apo_e4",nrow(modulespecific.PCx.apo_e4))

modulespecific.combine<-rbind(modulespecific.TCx.Disease_state,modulespecific.HIP.Disease_state,modulespecific.TCx.age,modulespecific.HIP.age,modulespecific.PCx.age,modulespecific.TCx.apo_e4,modulespecific.PCx.apo_e4)

age_disease.TCx$brain<-rep("TCx",nrow(age_disease.TCx))
age_disease.HIP$brain<-rep("HIP",nrow(age_disease.HIP))
age_disease.PCx$brain<-rep("PCx",nrow(age_disease.PCx))
age_disease.TCx$trait<-rep("age or disease",nrow(age_disease.TCx))
age_disease.HIP$trait<-rep("age or disease",nrow(age_disease.HIP))
age_disease.PCx$trait<-rep("age or disease",nrow(age_disease.PCx))

age_disease.combine<-rbind(age_disease.TCx,age_disease.HIP,age_disease.PCx)

disease.TCx$brain<-rep("TCx",nrow(disease.TCx))
disease.HIP$brain<-rep("HIP",nrow(disease.HIP))
disease.TCx$trait<-rep("disease",nrow(disease.TCx))
disease.HIP$trait<-rep("disease",nrow(disease.HIP))

disease.combine<-rbind(disease.TCx,disease.HIP)

outmodule.list<-list()
outmodule.list$modulespecific<-modulespecific.combine
outmodule.list$age_or_disease<-age_disease.combine
outmodule.list$disease<-disease.combine
write.xlsx(outmodule.list,paste(outPath,"modulespecific_age-or-disease_disease_module-list_all.xlsx",sep = '/'))
```

################################################################################
#
# Enrichment analysis of testing genesets in WGCNA modules
#
################################################################################

## load data

```{r}
# consensusAnalysis-CombinedNetworkResults

consensus_network.res<-read.csv("./results/consensusAnalysis-CombinedNetworkResults.csv")

# background gene sets

burdenresult_dementia<-read.csv("./metadata/burdenresult_dementia_annotatedforLR.csv") %>% dplyr::select(c("entrez_id","ensembl_gene_id","GENE","transcript_length","percentage_gene_gc_content","P_DOM_adjustFDR"))
burdenresult_AD<-read.csv("./metadata/burdenresult_AD_annotatedforLR.csv") %>% dplyr::select(c("entrez_id","ensembl_gene_id","GENE","transcript_length","percentage_gene_gc_content","P_DOM_adjustFDR"))
burdenresult_earlyAD<-read.csv("./metadata/burdenresult_earlyAD_annotatedforLR.csv") %>% dplyr::select(c("entrez_id","ensembl_gene_id","GENE","transcript_length","percentage_gene_gc_content","P_DOM_adjustFDR"))

colnames(burdenresult_dementia)<-c("entrezgene_id",colnames(burdenresult_dementia)[colnames(burdenresult_dementia) %!in% "entrez_id"])
colnames(burdenresult_AD)<-c("entrezgene_id",colnames(burdenresult_AD)[colnames(burdenresult_AD) %!in% "entrez_id"])
colnames(burdenresult_earlyAD)<-c("entrezgene_id",colnames(burdenresult_earlyAD)[colnames(burdenresult_earlyAD) %!in% "entrez_id"])

burdenresult_dementia$disease<-rep(0,nrow(burdenresult_dementia))
burdenresult_AD$disease<-rep(0,nrow(burdenresult_AD))
burdenresult_earlyAD$disease<-rep(0,nrow(burdenresult_earlyAD))

# testing gene sets

dementia_PPI_cluster<-read.table("./metadata/dementia_PPI_cluster.txt")[,1]
AD_PPI_cluster<-read.table("./metadata/AD_PPI_cluster.txt")[,1]
earlyAD_PPI_cluster<-read.table("./metadata/ADearly_PPI_cluster.txt")[,1]

list_burden_dementia<-read.table("./metadata/burdengenes_dementia_list.txt")[,1]
list_burden_AD<-read.table("./metadata/burdengenes_AD_list.txt")[,1]
list_burden_earlyAD<-read.table("./metadata/burdengenes_ADearly_list.txt")[,1]

list_caseonly_dementia<-read.table("./metadata/list_TRAPD_caseonly_dementia_entrezid.txt")[,1]
list_caseonly_AD<-read.table("./metadata/list_TRAPD_caseonly_AD_entrezid.txt")[,1]
list_caseonly_earlyAD<-read.table("./metadata/list_TRAPD_caseonly_earlyAD_entrezid.txt")[,1]
list_caseonly_and_burden_all<-read.table("./metadata/list_TRAPD_caseonly_mutated_an_burdengenes_entrezid.txt")[,1]

list_case_AD<-read.table("./metadata/list_case_AD_entrezid.txt")[,1]
list_case_earlyAD<-read.table("./metadata/list_case_earlyAD_entrezid.txt")[,1]
list_case_dementia<-read.table("./metadata/list_case_dementia_entrezid.txt")[,1]
```

## Section 1: Prepare background annotation data

```{r, message=FALSE, warning=FALSE}
WGCNAmodule.anno_dementia<-get_WGCNAmodule_annotation(geneinfo = consensus_network.res, df.background = burdenresult_dementia)
WGCNAmodule.anno_AD<-get_WGCNAmodule_annotation(geneinfo = consensus_network.res, df.background = burdenresult_AD)
WGCNAmodule.anno_EOAD<-get_WGCNAmodule_annotation(geneinfo = consensus_network.res, df.background = burdenresult_earlyAD)
```

## Section 2: module enrichment test

```{r, message=FALSE, warning=FALSE}
module_enrich_res.burdendementia<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_dementia,disease_genes = list_burden_dementia)
module_enrich_res.burdenAD<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_AD,disease_genes = list_burden_AD)
module_enrich_res.burdenearlyAD<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_EOAD,disease_genes = list_burden_earlyAD)

module_enrich_res.caseonly_dementia<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_dementia,disease_genes = list_caseonly_dementia)
module_enrich_res.caseonly_AD<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_AD,disease_genes = list_caseonly_AD)
module_enrich_res.caseonly_earlyAD<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_EOAD,disease_genes = list_caseonly_earlyAD)

module_enrich_res.case_dementia<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_dementia,disease_genes = list_case_dementia)
module_enrich_res.case_AD<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_AD,disease_genes = list_case_AD)
module_enrich_res.case_earlyAD<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_EOAD,disease_genes = list_case_earlyAD)

module_enrich_res.PPI_dementia<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_dementia,disease_genes = dementia_PPI_cluster)
module_enrich_res.PPI_AD<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_AD,disease_genes = AD_PPI_cluster)
module_enrich_res.PPI_earlyAD<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_EOAD,disease_genes = earlyAD_PPI_cluster)
```

## Section 3: filter significant results

```{r, message=FALSE, warning=FALSE}
sig_WGCNAmodule.burdendementia<-get_WGCNAmodule_enrichfilter(module_enrich_res.burdendementia,p_adjust_value = 0.1)
sig_WGCNAmodule.burdenAD<-get_WGCNAmodule_enrichfilter(module_enrich_res.burdenAD,p_adjust_value = 0.1)
sig_WGCNAmodule.burdenearlyAD<-get_WGCNAmodule_enrichfilter(module_enrich_res.burdenearlyAD,p_adjust_value = 0.1)

sig_WGCNAmodule.caseonly_dementia<-get_WGCNAmodule_enrichfilter(module_enrich_res.caseonly_dementia,p_adjust_value = 0.1)
sig_WGCNAmodule.caseonly_AD<-get_WGCNAmodule_enrichfilter(module_enrich_res.caseonly_AD,p_adjust_value = 0.1)
sig_WGCNAmodule.caseonly_earlyAD<-get_WGCNAmodule_enrichfilter(module_enrich_res.caseonly_earlyAD,p_adjust_value = 0.1)

sig_WGCNAmodule.case_dementia<-get_WGCNAmodule_enrichfilter(module_enrich_res.case_dementia,p_adjust_value = 0.1)
sig_WGCNAmodule.case_AD<-get_WGCNAmodule_enrichfilter(module_enrich_res.case_AD,p_adjust_value = 0.1)
sig_WGCNAmodule.case_earlyAD<-get_WGCNAmodule_enrichfilter(module_enrich_res.case_earlyAD,p_adjust_value = 0.1)

sig_WGCNAmodule.PPI_dementia<-get_WGCNAmodule_enrichfilter(module_enrich_res.PPI_dementia,p_adjust_value = 0.1)
sig_WGCNAmodule.PPI_AD<-get_WGCNAmodule_enrichfilter(module_enrich_res.PPI_AD,p_adjust_value = 0.1)
sig_WGCNAmodule.PPI_earlyAD<-get_WGCNAmodule_enrichfilter(module_enrich_res.PPI_earlyAD,p_adjust_value = 0.1)
```

### save significant enrichment results

```{r}
xl_list_module<-list()

xl_list_module$sig_burden_dementia<-sig_WGCNAmodule.burdendementia
xl_list_module$sig_burden_AD<-sig_WGCNAmodule.burdenAD
xl_list_module$sig_burden_earlyAD<-sig_WGCNAmodule.burdenearlyAD

xl_list_module$sig_caseonly_dementia<-sig_WGCNAmodule.caseonly_dementia
xl_list_module$sig_caseonly_AD<-sig_WGCNAmodule.caseonly_AD
xl_list_module$sig_caseonly_earlyAD<-sig_WGCNAmodule.caseonly_earlyAD

xl_list_module$sig_case_dementia<-sig_WGCNAmodule.case_dementia
xl_list_module$sig_case_AD<-sig_WGCNAmodule.case_AD
xl_list_module$sig_case_earlyAD<-sig_WGCNAmodule.case_earlyAD

xl_list_module$sig_PPI_dementia<-sig_WGCNAmodule.PPI_dementia
xl_list_module$sig_PPI_AD<-sig_WGCNAmodule.PPI_AD
xl_list_module$sig_PPI_earlyAD<-sig_WGCNAmodule.PPI_earlyAD

## output result tables

write.xlsx(xl_list_module,"./results/WGCNAmodule_enrich_sig.xlsx")
```

################################################################################
#
# Plot drawing
#
################################################################################

## prepare data for visualization

```{r message=FALSE, warning=FALSE}
## collect all the significant modules

ls.sig_WGCNAmodule<-unique(c(sig_WGCNAmodule.burdenAD$ModuleColor,sig_WGCNAmodule.caseonly_dementia$ModuleColor,sig_WGCNAmodule.caseonly_AD$ModuleColor,sig_WGCNAmodule.PPI_dementia$ModuleColor,sig_WGCNAmodule.PPI_earlyAD$ModuleColor)) %>% sort

## add group and combine

plt.burdendementia<-module_enrich_res.burdendementia %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("burdened dementia",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)
plt.burdenAD<-module_enrich_res.burdenAD %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("burdened AD",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)
plt.burdenearlyAD<-module_enrich_res.burdenearlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("burdened EOAD",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)

plt.caseonly_dementia<-module_enrich_res.caseonly_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("case-only dementia",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)
plt.caseonly_AD<-module_enrich_res.caseonly_AD %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("case-only AD",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)
plt.caseonly_earlyAD<-module_enrich_res.caseonly_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("case-only EOAD",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)

plt.case_dementia<-module_enrich_res.case_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("case dementia",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)
plt.case_AD<-module_enrich_res.case_AD %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("case AD",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)
plt.case_earlyAD<-module_enrich_res.case_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("case EOAD",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)

plt.PPI_dementia<-module_enrich_res.PPI_dementia %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("PPI dementia",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)
plt.PPI_AD<-module_enrich_res.PPI_AD %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("PPI AD",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)
plt.PPI_earlyAD<-module_enrich_res.PPI_earlyAD %>% dplyr::filter(rownames(.) %in% ls.sig_WGCNAmodule) %>% cbind(.,group=rep("PPI EOAD",length(ls.sig_WGCNAmodule))) %>% arrange(ModuleColor)

# res enrich plt combine

plt.all<-rbind(plt.burdendementia,plt.burdenAD,plt.burdenearlyAD,plt.caseonly_dementia,plt.caseonly_AD,plt.caseonly_earlyAD,plt.case_dementia,plt.case_AD,plt.case_earlyAD,plt.PPI_dementia,plt.PPI_AD,plt.PPI_earlyAD)

# output significant result tables

write.xlsx(plt.all,"./results/WGCNAmodule_sigvisual_table_allgroups.xlsx")
```

## data visualization

### plot drawing

```{r, message=FALSE, warning=FALSE}
heatmap.plt.all<-get_heatmatrix(WGCNAmodule_enrich=plt.all,ls.sig_WGCNAmodule,testing_set_name="WGCNA module enrichment")
```

### output the plots

```{r}
# define trellis object saving function

save_trellis <- function(heatmap_plot,ls.sig_WGCNAmodule,file_name,save_path){
 pdf(paste(save_path,file_name,sep = '/'),height=20,width=2+20*length(ls.sig_WGCNAmodule)/12)
 print(heatmap_plot)
 dev.off()
}

# save plot

save_trellis(heatmap.plt.all,ls.sig_WGCNAmodule,"Plots/plot_WGCNAmodule_enrichment_all.pdf",outPath)
```

################################################################################
#
# GO BP analysis of module genes
#
################################################################################

## GO BP enrichment test

### clusterprofiler

```{r, message=FALSE, warning=FALSE}
module_pathway_clusterprofiler <- get_module_pathway_clusterprofiler(module_list = ls.sig_WGCNAmodule,geneinfo = consensus_network.res)
write.xlsx(module_pathway_clusterprofiler,paste(outPath,"/GOBP_WGCNA_sigmodule_clusterprofiler.xlsx",sep = ''),rowNames = F)

unique(module_pathway_clusterprofiler$ModuleColor)
```

## module enrichment plot

### overlap of sig_WGCNAmodule and trait-associated modules

```{r}
ME.age_disease<-age_disease.traits$MEColorNames
ME.modulespecific<-modulespecific.traits$MEColorNames
ME.sig_WGCNAmodule.clusterprofiler<-unique(module_pathway_clusterprofiler$ModuleColor)

input.sig.age_disease<-list(ME.age_disease,ME.sig_WGCNAmodule.clusterprofiler)
input.sig.modulespecific<-list(ME.modulespecific,ME.sig_WGCNAmodule.clusterprofiler)

output.sig.age_disease<-attr(venn(input.sig.age_disease), "intersections")
output.sig.modulespecific<-attr(venn(input.sig.modulespecific), "intersections")

overlap.sig.age_disease<-output.sig.age_disease$`A:B`
overlap.sig.modulespecific<-output.sig.modulespecific$`A:B`
```

### plot the overlap modules

```{r}
# GO:BP barplot

module_list<-unique(c(overlap.sig.age_disease,overlap.sig.modulespecific))
term_num<-rep(10,length(module_list))
overlap.plt<-get_pathway_barplot(module_pathway_clusterprofiler,module.list = module_list,term_num.list = term_num)

# output plots

pdf(paste(outPath,"Plots/plot_GOBP_moduleoverlap_age-or-disease-with-trait-specific_sigWGCNAtest.pdf",sep = '/'))
for(i in 1:length(overlap.plt)){
  print(overlap.plt[[i]])
}
dev.off()
```

## add evidence to the ranking list

### in age-or-disease related modules or not

```{r}
# the previous script in "spatiotemporal_enrichment_fractionandADknown"

evidence_df<-read.xlsx("./metadata/evidence-ranking_table_all-case-only-and-burdened-genes.xlsx")
gene_list<-consensus_network.res %>% dplyr::filter(ModuleColor %in% age_disease.traits$MEColorNames) %>% dplyr::pull(EntrezID)
evidence_df$age_disease_modules<-evidence_df$ENTREZID %>% lapply(.,function(x){if(x %in% gene_list){1}else{0}}) %>% unlist
```

### co-occur with AD known genes in the same modules or not

```{r}
# AD known gene exploration in WGCNA results

## overlap

AD_known_genes<-read.table("./metadata/AD_knownsig_geneset_v2.txt")[,1]
AD_known_genes.overlapmodules<-consensus_network.res %>% dplyr::filter(EntrezID %in% AD_known_genes) %>% dplyr::pull(ModuleColor) %>% unique %>% lapply(.,function(x){if(x != "grey"){x}}) %>% unlist
AD_known_genes.overlapmodules.df<-consensus_network.res %>% dplyr::filter(EntrezID %in% AD_known_genes) %>% dplyr::pull(ModuleColor) %>% lapply(.,function(x){if(x != "grey"){x}}) %>% unlist %>% table %>% as.data.frame %>% arrange(desc(Freq))
AD_known_genes.overlapmodules.genes<-consensus_network.res %>% dplyr::filter(EntrezID %in% AD_known_genes) %>% dplyr::select(EntrezID,ModuleColor) %>% dplyr::filter(ModuleColor!="grey") %>% arrange(ModuleColor)

## module enrichment

module_enrich_res.AD_known_genes<-get_WGCNAmodule_Enrichment(WGCNAmodule.anno_dementia,disease_genes = AD_known_genes)
sig_WGCNAmodule.AD_known_genes<-get_WGCNAmodule_enrichfilter(module_enrich_res.AD_known_genes,p_adjust_value = 0.1)

## output results of AD known gene exploration

AD_known_genes_res.list<-list()
AD_known_genes_res.list$overlapmodules<-AD_known_genes.overlapmodules.genes
AD_known_genes_res.list$sig_WGCNAmodule<-sig_WGCNAmodule.AD_known_genes
write.xlsx(AD_known_genes_res.list,paste(outPath,"AD-known-genes_overlapped-and-sigenriched_modules.xlsx",sep = '/'))

# add evidence: co-occur with AD known genes in the same modules

gene_list<-consensus_network.res %>% dplyr::filter(ModuleColor %in% AD_known_genes.overlapmodules) %>% dplyr::pull(EntrezID)
evidence_df$overlap_ADknowngenes_modules<-evidence_df$ENTREZID %>% lapply(.,function(x){if(x %in% gene_list){1}else{0}}) %>% unlist
```

### in cilium or microtube-related modules or not

```{r}
# mark related modules

mark<-module_pathway_clusterprofiler$Description %>% grepl("sperm",.) %>% lapply(.,isFALSE) %>% unlist
filter.module_pathway_clusterprofiler<-module_pathway_clusterprofiler %>% filter(mark)
mark<-grepl("microtubule|axoneme|cytoskeleton|cilium",filter.module_pathway_clusterprofiler$Description)
filter.module_pathway_clusterprofiler %<>% filter(mark)
write.xlsx(filter.module_pathway_clusterprofiler,paste(outPath,"allsig_cilium-or-microtube-related-modules_pathway-clusterprofiler.xlsx",sep = '/'))
filter.module<-filter.module_pathway_clusterprofiler %>% dplyr::pull(ModuleColor) %>% unique

# add evidence

gene_list<-consensus_network.res %>% dplyr::filter(ModuleColor %in% filter.module) %>% dplyr::pull(EntrezID)
evidence_df$microtubule_cilium_related_modules<-evidence_df$ENTREZID %>% lapply(.,function(x){if(x %in% gene_list){1}else{0}}) %>% unlist

write.xlsx(evidence_df,"./metadata/evidence-ranking_table_all-case-only-and-burdened-genes.xlsx")

# the next script is "singlecell_marker_enrichment_analysis_CellMarker"
```

## save image

```{r}
save.image("./WGCNA_brainregioncombined_enrichment_analysis-GO.RData")
```
